### 2024-10-21 main_widom_morse.tex
first commit

\documentclass[12pt,a4paper]{article}

\usepackage[english]{babel}
\usepackage[T2A]{fontenc}
\usepackage[cp1251]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{color}
\usepackage{amsfonts}
\usepackage{wrapfig}
\usepackage{caption}

\usepackage{hyperref}
\usepackage{physics}

\textheight=26cm
\textwidth=17cm
\hoffset=0pt
\voffset=0cm
\oddsidemargin=0cm
\topmargin=-60pt
\footskip=30pt
\renewcommand{\baselinestretch}{1}

\newcommand{\no}{\nonumber}
\newcommand{\non}{\nonumber \\}
\newcommand{\ve}[1]{{\bf #1}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\sli}{\sum\limits}
\newcommand{\ili}{\int\limits}
\newcommand{\lp}{\left (}
\newcommand{\rp}{\right )}
\newcommand{\lb}{\left [}
\newcommand{\rb}{\right ]}
\newcommand{\lbr}{\left \{}
\newcommand{\rbr}{\right \}}
\newcommand{\ld}{\left .}
\newcommand{\rd}{\right .}
\newcommand{\vk}{\vec{k}}
\newcommand{\vl}{\vec{l}}
\newcommand{\cB}{{\cal B}}
\newcommand{\cM}{{\frak M}}
\newcommand{\rhok}{\rho_{\vec{k}}}
\newcommand{\rhomk}{\rho_{-\vec{k}}}
\newcommand{\etak}{\eta_{\vec{k}}}
\newcommand{\etamk}{\eta_{-\vec{k}}}

\begin{document}

\binoppenalty=10000
\relpenalty=10000
\begin{center}
	\renewcommand{\theequation}{\arabic{section}}
	\section*{Crossover line in the supercritical cell fluid model (I haven't thought up a good title yet)}
	\setcounter{equation}{0}
\end{center}

\subsection*{Abstract}

\section{Introduction}
A supercritical fluid is a state of matter that occurs when a substance is subjected to temperatures and pressures above its critical point. Below the critical point the substance exhibits a liquid-gas coexistence. At the critical point, the distinction between the liquid and gas phases of a substance disappears, and beyond this point, the substance exists in a single continuous supercritical state. Despite the absence of first-order phase transitions above the critical point, fluids exhibit distinct liquid-like and gas-like structures under supercritical conditions, and display a unique mix of properties. With densities comparable to liquids they possess low viscosity and high diffusivity like gases resulting in high solvent power. Moreover, the supercritical properties are instantly tunable by changing pressure and temperature. This is the reason why supercritical fluids have diverse applications across various industries \cite{KnezEtAl2014}.

In the context of supercritical fluids, the Widom line, Frenkel line, and percolation line are theoretical and physical constructs used to describe the complex behavior of fluids beyond the critical point.
All three concepts provide a way to describe how fluids behave in the supercritical region, where the traditional liquid-gas phase distinction no longer applies.
The Frenkel line \cite{BrykEtAl2017} focuses on the dynamics of molecular motion, separating regimes where the fluid behaves more like a liquid (with oscillatory molecular motion) from those where it behaves more like a gas (with only diffusive motion). The percolation line \cite{CampiEtAl2001} is related to spatial structures, marking the threshold where clusters of fluid or density fluctuations start forming a continuous, system-spanning network.

In the present work we are particularly interested in the Widom line \cite{XuEtAl2005,McMillanStanley2010,SimeoniEtAl2010} which focuses on thermodynamic properties and represents a conditional boundary between liquid-like and gas-like regions in the supercritical fluid, based on response functions. Thus the Widom line \--- the locus of correlation length maxima emanating from the critical point \cite{XuEtAl2005}, is treated as the continuation of the coexistence line above the critical point.   
Some thermodynamic response functions, for example, isobaric heat capacity and isothermal compressibility, also exhibit maxima in the supercritical region. In the critical region beyond the critical point the maxima lines of the correlation length and response functions virtually merge into a single line \cite{ArtemenkoEtAl2017} and then diverge from each other as the temperature rise. Thus in the critical region, the Widom line is viewed as the maxima locus of thermodynamic response functions \cite{Proctor2020Book}.

% directly from the equation of state 
[Here will be a small paragraph about challenges on developing the global equation of state directly predicting crossover line in supercritical fluids]
% we would like to share an idea

In Section~\ref{sec:eos}, we present the modified Morse potential used in our work, the explicit expression for the equation of state, and plot of pressure isotherms at temperatures below and above the critical one. In Section~\ref{sec:widom}, we display the Widom line as the maxima locus of the isothermal compressibility and introduce the supercritical crossover line obtained directly from the equation of state of the cell fluid model.

\section{\label{sec:eos}The equation of state: pressure as a function of temperature and chemical potential}

In the present work we take advantage of the results obtained in \cite{KozlovskiiDobush2020,DobushEtAl2024}. In \cite{KozlovskiiDobush2020}, the phase space of collective variables%\footnote{The term "collective variables" is commonly used to refer to a specific category of variables unique to each physical system. In the context of liquid-vapour systems, these collective variables are linked to modes of density deviation from the critical density. This set of collective variables encompasses those which average values are connected to order parameters.Thus, the depiction of the phase transition within the phase space of collective variables appears compelling.} 
is used for calculating the grand partition function of a cell fluid model in the zero-mode approximation and consequently deriving the explicit forms of the equation of state (EoS) both in pressure-temperature-chemical potential terms and pressure-temperature-density terms. The resulting EoSs are applicable in a wide range of temperatures below and above the critical point. In \cite{DobushEtAl2024}, we used these equations as the basis for the calculation and graphical representation of the thermodynamic response functions (isothermal compressibility, thermal pressure coefficient, thermal expansion) of the cell fluid model with modified Morse potential in the supercritical region.

**************************************************************

[The following is the literal copy (compilation) from our recent preprint on response functions. Perhaps we should rephrase it, or change it somehow. Please help.]

**************************************************************

The potential of interaction between particles is taken in a form of a modified Morse potential
\begin{eqnarray}
	\label{def:mod_morse}
	U(r) & = & \varepsilon C_{H} \left[A {\rm e}^{-n_0(r-R_0)/\alpha} \right.
	\nonumber\\
	&& + \left. {\rm e}^{-\gamma(r - R_0)/\alpha} -2{\rm e}^{-(r-R_0)/\alpha}\right],
\end{eqnarray}
where $R_0$ is the coordinate of the potential minimum, $\alpha$ is an effective range of interaction, $\gamma$ and $n_0$ are parameters of the model. Other two constants $C_{H}$ and $A$ are expressed via $\gamma$ and $n_0$ as follows
\begin{equation}
	C_{H} = \frac{n_0}{n_0 + \gamma - 2}, \quad A = \frac{2 - \gamma}{n_0},
\end{equation}
where $\varepsilon$ is the depth of the potential well at $r=R_0$. Applying $\gamma=2$ reduce the  to the ordinary Morse potential~\cite{Morse1929} at $\gamma=2$. For a more detailed discussion of such modified Morse potential, see Sections~1 and~2 in~\cite{KozlovskiiDobush2020}.

Our modification contains an additional repulsive term, which enables us to single out a reference system (in the reciprocal space) and apply the method of collective variables to calculating the grand partition function~\cite{KozlovskiiDobush2016}.

%The explicit form of the equation of state in terms of chemical potential-temperature in the case of $T\geq T_c$ is as follows:

The equation of state obtained in~\cite{KozlovskiiDobush2020} reads
\begin{equation}
	\label{eq:eosMT}
	Pv\beta = E_\mu(M, T) + M \bar \rho_0 + \frac{1}{2} d \bar \rho_0^2 - \frac{a_4}{24} \bar \rho_0^4.
\end{equation}
The quantities in the left-hand side of the equation are $P$, the pressure; $\beta = (k_{\rm B} T)^{-1}$, the inverse temperature; $k_{\rm B}$, the Boltzmann constant; $T$, the temperature; $v$, cell volume. The quantities in the right-hand side are, in general, functions of the temperature $T$ and the chemical potential $\mu$. Let us present their expressions explicitly.

First, the quantity $M$ depends linearly on the chemical potential
\begin{align}\label{chem_pot}
	&	M = \frac{\tilde\mu}{W(0)} + g_1 - \frac{g_3}{g_4} d - \frac{1}{6} \frac{g_3^3}{g_4^2}, \\
	&	\tilde\mu=\mu-\mu_0(1+\tau),
\end{align}
where $\mu_0$ is some positive constant, $\tau$ is the relative temperature $\tau = (T - T_c) / T_c$, $T_c$ is the critical temperature. We will call $M$ the effective chemical potential.

The quantity $W(0)$ is expressed via parameters of the potential~(\ref{def:mod_morse}) as follows
\begin{equation}
	W(0) = \Phi^{(r)}(0) \left[ B - 1 + \chi_0 + \tau (\chi_0 + A_\gamma) \right],
\end{equation}
where
\begin{align*} 
	& B = 2 \gamma^3 e^{(1-\gamma)R_0/\alpha},
	\nonumber \\
	& A_\gamma = A e^{(n_0-\gamma)R_0/\alpha} \left( \gamma / n_0\right)^3, 
\end{align*}
and $\Phi^{(r)}(0)$ is the Fourier transform of the repulsive part of the potential at $\abs{\vb k}=0$
\begin{equation*}
	\Phi^{(r)}(0) = \varepsilon C_H 8\pi {\rm e}^{\gamma R_0/\alpha} \left(\frac{\alpha}{\gamma R_0}\right)^3.
\end{equation*}

The parameter $\chi_0$ is used in~\cite{KozlovskiiDobush2020} to single out a contribution in the Fourier transform of the potential that is treated as a reference system defined in the reciprocal space, and is selected as $\chi_0 = 0.07$~\cite{KozlovskiiDobush2020}.

The coefficients $g_n$ are given by the formulas:
\begin{align}
	& g_0 = \ln T_0, \qquad g_1 = T_1/T_0, \qquad g_2 = T_2/T_0 - g_1^2,  
	\nonumber \\
	& g_3 = T_3/T_0 - g_1^3 - 3g_1 g_2, 
	\nonumber\\
	& g_4 = T_4/T_0 - g_1^4 - 6 g_1^2 g_2 - 4 g_1 g_3 - 3 g_2^2, 
	\nonumber\\
	& a_4 = -g_4,
\end{align}
where $T_n(p,\alpha^*)$ are the following special functions
\begin{equation}
	T_n(p,\alpha^*) = \sum_{m=0}^{\infty} \frac{(\alpha^*)^m}{m!} m^n {\rm e}^{-pm^2}.
\end{equation}
Here $\alpha^*=v e^{\beta_c\mu_0}$, and the parameter $p$ has the form
\begin{equation}
	p = \frac{\beta_c}{2} \Phi^{(r)}(0) [\chi_0 + A_\gamma].
\end{equation}

The quantity $\beta_c$ denotes the critical value of the inverse temperature. In~\cite{KozlovskiiDobush2020} it was found that
\begin{equation*}
	\varepsilon\beta_c = 0.200, \quad \frac{k_{\rm B} T_c}{\varepsilon} = 4.995.
\end{equation*}
We also use the reduced temperature defined as $T^* = k_{\rm B}T/\varepsilon$, and thus its critical value $T^*_c = 4.995.$

Since $p$ is independent of temperature, the coefficients $g_n$ are also independent of temperature. The numerical values for other coefficients used in this paper are the same as those in~\cite[eqs.~(5), (23), and~(24)]{KozlovskiiDobush2020}:
\begin{eqnarray}
	\label{params}
	\chi_0 = 0.07, & \quad \gamma = 1.65, \nonumber\\
	n_0 = 1.521, & \quad R_0/\alpha = 2.9544, \nonumber\\
	\alpha^* = 5.0 & \quad p = 1.0.
\end{eqnarray}

The quantity $d$ entering equations~\eqref{eq:eosMT} and~\eqref{chem_pot} is a function of temperature
\begin{equation}
	\label{def:D0}
	d = g_2 - \frac{1}{2} \frac{g_3^2}{g_4} - \frac{1}{\beta W(0)}.
\end{equation}
The condition $d = 0$ defines the critical temperature~\cite{KozlovskiiDobush2020}
\begin{equation}
	k_{\rm B}T_c = \left(g_2 - \frac{1}{2} \frac{g_3^2}{g_4} \right) (B - 1 + \chi_0) \Phi^{(r)}(0).
\end{equation}

The function $E_\mu(M, T)$ from the equation \eqref{eq:eosMT} is provided by
\begin{eqnarray}\label{eq:E_mu}
	E_\mu (M, T) & = & - \frac{\ln (2\pi \beta W(0))}{2 N_v}  
	\nonumber\\
	&& +  g_0 - \frac{\beta W(0)}{2} 
	\left(\frac{\tilde\mu}{W(0)} \right)^{2} 
	\nonumber\\
	&& - \frac{g_3}{g_4} {M} \! - \frac{g_3^2}{2 g_4^2}  d - \frac{1}{24} \frac{g_3^4}{g_4^3}. 
\end{eqnarray}
Here the quantity $N_v$ defines the number of cubic cells in volume $V$ for the initial model.
In the thermodynamic limit, $N_v \to \infty$, and thus, the first term can be neglected. The term $\tilde{\mu}/W(0)$ can be expressed in terms of $M$ using~\eqref{chem_pot}. The temperature and the inverse temperature can always be expressed in terms of the reduced temperature and a corresponding critical value:
\begin{equation*}
	T = T_c(1+\tau), \quad \beta = \beta_c (1 + \tau)^{-1}.
\end{equation*}

The quantity $\bar{\rho}_0$ is a solution to the following cubic equation
\begin{equation}\label{eq:ro_M}
	M + d \bar\rho_0 - \frac{a_4}{6} \bar\rho_0^3 = 0.
\end{equation}
For any $\tau > 0$, the latter equation has one real root
\begin{equation}\label{eq:ro_MT}
	\bar \rho_0 = \left(- \frac{3 M}{g_4} + \sqrt{Q_t}\right)^{1/3} - \left(  \frac{3 M}{g_4} + \sqrt{Q_t} \right)^{1/3},
\end{equation}
where
\begin{equation}
	Q_t = \left(  \frac{2d}{g_4}\right)^3 + \left( -\frac{3 M}{g_4}\right)^2, \qquad g_4<0.
\end{equation}
Thus, $\bar{\rho}_0$ is a function of the temperature and the chemical potential.

Let us introduce the reduced pressure
\begin{equation}
	\label{def:reduced_pres}
	P^* \equiv \frac{P v}{\varepsilon}.
\end{equation}
Considering the equation of state~\eqref{eq:eosMT}, $P^*$ is explicitly written as
\begin{eqnarray}
	\label{eq:eosPTM_reduced}
	P^* & = & (1 + \tau)T^*_c \bigg[ E_\mu(M, T) 
	\nonumber\\
	&& + M \bar \rho_0 + \frac{1}{2} d \bar \rho_0^2 - \frac{a_4}{24} \bar \rho_0^4
	\bigg].
\end{eqnarray}

*****************************************************************************************

End of literal copy

*****************************************************************************************

\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{f1.pdf} 
	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the effective chemical potential $M$ at $T \geq T_c$ represented by black lines. Grey lines correspond to isotherms of pressure at $T < T_c$ based on the results for the same model taken from~\cite{KozlovskiiDobush2020}. 
	}\label{fig1}
\end{figure}
Figure~\ref{fig1} illustrates the relationship between the reduced pressure $P^*$ and the effective chemical potential $M$ for various relative temperature values $\tau$. Below the critical temperature ($\tau < 0$), іsotherms are curves consisting of two segments connected by a vertex (grey lines in Figure~\ref{fig1}). This form indicates a first-order phase transition occurs in the system. Within our approach~\cite{KozlovskiiDobush2020} at $T<T_c$, negative values of the effective chemical potential mean the system is in a gaseous state. The vertex $M=0$ is a point of phase coexistence. Positive values of the effective chemical potential correspond to the liquid state. In both cases, the pressure tends to increase with $M$, but the slope of the liquid isotherm is much sharper. At the critical temperature pressure is a smooth plane curve (black dotted line in Figure~\ref{fig1}) with an inflection point at $M=0$. Thus, at the critical point $M=0, \tau=0$, $P^*_c = 1.606$. Beyond the critical temperature, the pressure behaves as a monotonically increasing function, and no phase transitions are observed. Summing up, since below the critical temperature, $M=0$ stands for phase coexistence, we obtain a phase diagram in pressure-temperature terms (see Figure~\ref{fig2}) by simply using this condition in the equation of state \eqref{eq:eosPTM_reduced}. 
\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{f2.pdf} 
	\vskip-3mm\caption{Phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse} based on the equation of state~\eqref{eq:eosPTM_reduced}. 
	}\label{fig2}
\end{figure}

%\section{Ентропія}

\section{\label{sec:widom}Crossover line between gas-like and liquid-like structures of the supercritical fluid. The Widom line}

 In the supercritical region, the substance no longer behaves distinctly as either a liquid or a gas. Instead, it exists as a single, continuous phase with no sharp liquid-gas boundary. Nevertheless, fluids exhibit distinct liquid-like and gas-like structures beyond the critical temperature. A way to represent a conditional boundary between liquid-like and gas-like regions in the supercritical fluid is the Widom line \--- a continuation of the coexistence line based on the maxima locus of some thermodynamic response functions~\cite{XuEtAl2005,McMillanStanley2010,SimeoniEtAl2010}. This happens because near the critical point, the system experiences very large fluctuations in density, and these fluctuations occur over increasingly large distances. Although, the equation of state is derived in the mean-field-like approximation,   in~\cite{DobushEtAl2024} we established that the isothermal compressibility of our cell fluid model diverge in the critical point, and consequently it display maxima in the supercritical region (see Figures~\ref{fig3}). It means the possibility to plot the Widom line, and to expand the phase diagram in Figure~\ref{fig2} beyond the critical point.
 In what follows we consider this in more detail.

The isothermal compressibility $\kappa_T$ is defined by
\begin{equation}
	\label{def:isotherm_compres}
	\kappa_T = -\frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_{T, N}.
\end{equation}
In \cite{DobushEtAl2024} we derived expression for the reduced isothermal compressibility $\kappa^*_T \equiv \dfrac{\varepsilon \kappa_T}{v}$  in terms of thermodynamic derivatives suitable for the reduced form of equation of state in pressure-temperture-chemical potential terms \eqref{eq:eosPTM_reduced} 
It is expressed in terms of the reduced pressure $P^*$ and the reduced particle number density $\rho^*$ as follows
\begin{eqnarray}
	\label{eq:kappa_star_m1}
	\kappa^*_T & = & \frac{\varepsilon}{{\rho^*}^2} \left(\frac{\partial \rho^*}{\partial \mu}\right)_T
	\\
	\label{eq:kappa_star_m}
	& = & \frac{\varepsilon^2}{{\rho^*}^2} \left(\frac{\partial^2 P^*}{\partial \mu^2}\right)_T.
\end{eqnarray}
The reduced particle number density $\rho^*$ is found by 
\begin{equation}
	\rho^* \equiv \frac{\langle N \rangle}{V} v = \left(\frac{\partial (Pv\beta)}{\partial (\beta \mu)}\right)_{T,V}.
\end{equation}
Taking explicit derivatives, we arrive at
\begin{equation}\label{eq:density}
	\rho^* = \rho^*_c - M + \frac{ \bar \rho_0}{\beta W(0)}.
\end{equation}
The quantity $\rho^*_c$ in the equation \eqref{eq:density} is the critical density~\cite{KozlovskiiDobush2020,DobushEtAl2024}
\begin{eqnarray}\label{eq:crit_dens}
	\rho^*_c & = &  g_1 - \frac{g_2 g_3}{g_4} + \frac{g_3^3}{3g_4^2}.
\end{eqnarray}
Its numerical value for parameters~\eqref{params} is $\rho^*_c = 0.978$.
Finally, derive the isothermal compressibility in explicit form  
\begin{eqnarray}
	\kappa_T^* & = & \frac{\epsilon}{\rho^{*2} W(0)} \left\{   \frac{Q_t^{-1/2}}{\beta W(0) g_4} \left[ \bar \rho_0 - 2 \left( - \frac{3M}{g_4} + \sqrt{Q_t} \right)^{\frac{1}{3}} \right] -1 \right\}.
\end{eqnarray}

**************************************************************************

[Figures only. Spots of text in progress.]

\begin{figure}[h!] 
	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
	\includegraphics[width=0.46\textwidth]{f3bb.pdf}
	%\includegraphics[width=\column]{f2b.pdf}
	%\includegraphics[width=\column]{f2c.pdf}
	\vskip-3mm
	\caption{The reduced isothermal compressibility $\kappa^*_T$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$. The two figures differ in the scale of $M$. Part~(a) covers a wider range of $M$. Part~(b) focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\kappa^*_T$.
	}
	\label{fig3}
\end{figure}

\begin{figure}[h!]
	\centering \includegraphics[width=0.5\textwidth]{f4.pdf}
	\vskip-3mm\caption{Complete phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse}. Comparison of the Widom line (maxima locus of isothermal compressibility) and the crossover line between gas-like and liquid-like structures of the supercritical cell fluid based on the equation of state~\eqref{eq:eosPTM_reduced}. An inset in the right bottom corner focuses on a range of $\tau$ around its critical value $-0.01 < \tau < 0.05$.	}\label{fig4}
\end{figure}

%\subsubsection{Лінія Відома (лінія максимумів стисливості)}

%\begin{figure}[h!]
%	\centering \includegraphics[width=0.5\textwidth]{f5.pdf}
%	\vskip-3mm\caption{Plot of the isothermal compressibility maxima line (the Widom line).	}\label{fig5}
%\end{figure}

%\subsubsection{Лінія розмежування газоподібної і рідиноподібної структур надкритичного плину, виходячи з рівняння стану при $M=0$}
%\begin{figure}[h!]
%	\centering \includegraphics[width=0.5\textwidth]{f6.pdf}
%	\vskip-3mm\caption{Plot of the crossover line between gas-like and liquid-like structures of the supercritical cell fluid based on the equation of state.	}\label{fig6}
%\end{figure}
%\subsubsection{Порівняння обох ліній}

\section{Conclusions}

[Spots of text in progress.]

\vskip3mm \textit{This work was supported by the National Research Foundation of Ukraine under the project No.~2023.03/0201.}

\begin{thebibliography}{99}

	%@incollection{KKD2018,
		%  title={Phase Transitions in a Continuum {C}urie-{W}eiss System: A Quantitative Analysis},
		%  author={Kozitsky, Yu. and Kozlovskii, M. and Dobush, O.},
		%  booktitle={Modern Problems of Molecular Physics},
		%  editor={Bulavin, L. A. Chalyi, A. V.},
		%  pages={229--251},
		%  year={2018},
		%  publisher={Springer},
		%  address={Cham, Switzerland}

		\bibitem{KnezEtAl2014} {\v{Z}}. Knez, E. Marko{\v{c}}i{\v{c}}, M. Leitgeb, M. Primo{\v{z}}i{\v{c}}, M. Knez Hrn{\v{c}}i{\v{c}}, M. {\v{S}}kerget, Industrial applications of supercritical fluids: A review, Energy, 77 (2014) 235-243. % https://doi.org/10.1016/j.energy.2014.07.044
		\bibitem{BrykEtAl2017} T. Bryk, F. A. Gorelli, I. Mryglod, G. Ruocco, M. Santoro, T. Scopigno. Behavior of Supercritical Fluids across the “Frenkel Line”. The Journal of Physical Chemistry Letters 2017 8 (20), 4995-5001 %	DOI: 10.1021/acs.jpclett.7b02176
		\bibitem{CampiEtAl2001} Campi, X., Krivine, H., Sator, N. Percolation line of self-bound clusters in supercritical fluids. Physica A: Statistical Mechanics and its Applications, 296(1-2), 24-30 (2001). %https://doi.org/10.1016/S0378-4371(01)00158-3
		\bibitem{XuEtAl2005}Xu, L., Kumar, P., Buldyrev, S. V., Chen, S.-H., Poole, P. H., Sciortino, F., Stanley, H. E. Relation between the Widom line and the dynamic crossover in systems with a liquid-liquid phase transition. Proceedings of the National Academy of Sciences, 102(46), 16558–16562  (2005). doi:10.1073/pnas.0507870102 
		\bibitem{McMillanStanley2010} McMillan, P., Stanley, H. Going supercritical. Nature Phys 6, 479–480 (2010). %https://doi.org/10.1038/nphys1711
		\bibitem{SimeoniEtAl2010} Simeoni, G., Bryk, T., Gorelli, F. et al. The Widom line as the crossover between liquid-like and gas-like behaviour in supercritical fluids. Nature Phys 6, 503–507 (2010). %https://doi.org/10.1038/nphys1683
		\bibitem{ArtemenkoEtAl2017} Artemenko, S., Krijgsman, P., Mazur, V. The Widom line for supercritical fluids. Journal of Molecular Liquids, 238, 122-128 (2017). %http://dx.doi.org/10.1016/j.molliq.2017.03.107
		\bibitem{Proctor2020Book} J.~E. Proctor, The liquid and supercritical fluid states of matter. CRC Press (2020). %[ISBN 9780429491443]

		\bibitem{KozlovskiiDobush2020}  M.P. Kozlovskii and O.A. Dobush,  Ukr. J. Phys. {\bf 65} 428 (2020).
		\bibitem{DobushEtAl2024} O.A. Dobush, M.P. Kozlovskii, R.V. Romanik, I.V. Pylyuk, arXiv preprint 2409.09786 (2024)
		\bibitem{KozlovskiiDobush2016}  M. Kozlovskii and O. Dobush, J. Mol. Liq., {\bf 215}, 58, (2016).

%		\bibitem{refb}  I.V. Pylyuk and O.A. Dobush,  Ukr. J. Phys. {\bf 65} 1080 (2020).
%		\bibitem{refc0} I.V. Pylyuk and M.P. Kozlovskii, O.A. Dobush, M.V. Dufanets, J. Mol. Liq., {\bf 355}, 122322, (2023).
%		\bibitem{refc1} I.V. Pylyuk and M.P. Kozlovskii, O.A. Dobush, Ukr. J. Phys., {\bf 68}, 601, (2023).
%		\bibitem{refc}  M.P. Kozlovskii and O.A. Dobush, Condens. Matter Phys., {\bf 20}, 23501,(2017).
%		\bibitem{refd}  M.P. Kozlovskii and I.V. Pylyuk, O.A. Dobush, Condens. Matter Phys., {\bf 21}, 43502, (2018).
%		\bibitem{ref8}  R.C. Lincoln and K.M. Koliwad, Phys. Rev., {\bf 157}, 463, (1967).
%		\bibitem{ref9}  J.K. Singh, J. Adhikari, and S.K. Kwak, Fluid Phase Equilibria, {\bf 248}, 1, (2006).

%		\bibitem{ref16} M.V.~Fedoryuk, \textit{Asymptotic methods in analysis} in {\it Analysis I: Integral Representations and Asymptotic Methods}, eds. M.A. Evgrafov and R.V. Gamkrelidze  (Encyclopaedia of Mathematical Sciences 13, Springer-Verlag Berlin Heidelberg, 1989) pp.~83--191.

%		\bibitem{ref11} A. J. Schultz and D. A. Kofke, Fluid Phase Equilibria, {\bf 409}, 12, (2016).
%		\bibitem{ref12} A. J. Masters, J. Phys.: Condens. Matter, {\bf 20}, 283102, (2008).
%		\bibitem{ref13} M. V. Ushcats, J. Chem. Phys., {\bf 140}, 234309, (2014).
%		\bibitem{ref14} D. Pini and G. Stell, N. B. Wilding, Mol. Phys., {\bf95}, 483, (1998).
%		\bibitem{ref15} C.-L. Lee and G. Stell, J. Hoye, J. Mol. Liq., {\bf 112}, 13, (2004).
%		\bibitem{ref1}  Y. Kozitsky and M. Kozlovskii, O. Dobush, In \emph{Modern Problems of Molecular Physics}, Springer, 229--251 (2018).
%		\bibitem{ref3}  M.P. Kozlovskii and O.A. Dobush, I.V. Pylyuk, Ukr. J. Phys. {\bf 62} 865 (2017).
%		\bibitem{ref4}  I. Yukhnovskii and V. Kolomiets, I. Idzyk, Condens. Matter Phys., {\bf 16}, 23604, (2013).
%		\bibitem{ref5}  I.R. Yukhnovskii, Condens. Matter Phys., {\bf 17}, 43001, (2014).

%		\bibitem{ref10} I.R. Yukhnovskii and M.P. Kozlovskii, I.V. Pylyuk, Phys. Rev. B, {\bf 66}, 134410, (2002).
%		\bibitem{ref17} D. Pini and G. Stell, N. B. Wilding, Mol. Phys., \textbf{95}, 483, (1998).
%		\bibitem{ref18} J.-M. Caillol, Mol. Phys., {\bf 104}, 1931, (2006).
%		\bibitem{ref19} I.R Yukhnovskii., Theoret. and Math. Phys., \textbf{194}, 224, (2018).
%		\bibitem{ref20} A. Oleinikova and L. Bulavin, V. Pipich, Chem. Phys. Let., {\bf 278}, 121, (1997).
%		\bibitem{ref21} G. Brunner, J. Supercrit. Fluids, {\bf 47}, 373, (2009).
%		\bibitem{ref22} G. Brunner, J. Supercrit. Fluids, {\bf 47}, 382, (2009).

%	\bibitem{ref1} 
%	\bibitem{ref2}
%	\bibitem{ref3}
%	\bibitem{ref4} 
%	\bibitem{ref5}
%	\bibitem{ref6}
%	\bibitem{ref7}
%	\bibitem{ref8}
%	\bibitem{ref9}
%	\bibitem{ref10}
%	\bibitem{ref11}
%	\bibitem{ref12}
%	\bibitem{ref13}
%	\bibitem{ref14}
%	\bibitem{ref15}
%	\bibitem{ref16}
%	\bibitem{ref17}

\end{thebibliography}

\end{document}

	The latter expression links the pressure $P$, the temperature $T$ ($\beta = 1/T$ is the inverse temperature), the cell's volume $v$ and the effective chemical potential $M$. In general $M$ is a function of the real chemical potential $\mu$:
\begin{align}
	&	M = \frac{\tilde\mu}{W(0)} + g_1 + n_c \tilde D(0) - \frac{1}{6} \frac{g_3^3}{g_4^2}, \\
	&	\tilde\mu=\mu-\mu^*(1+\tau),
\end{align}
here $\mu^*$ is some positive constant, and $\tau$ is the relative temperature $\tau = (T - T_c) / T_c$,
$T_c$ is the critical temperature. $W(0)$ is what we call an effective interaction potential\footnote{\textbf{explain} what is the effective interaction potential in a few words} at ${\bf k} = 0$:
\begin{equation}
	W(0) = \Phi^{(r)}(0) \left[ B - 1 + \chi_0 + \tau (\chi_0 + A_\gamma) \right].
\end{equation}
Here $\chi_0$ is a constant parameter, the quantities $B$ and $A_\gamma$ reads
\begin{align} 
	& B = 2 \gamma^3 e^{(1-\gamma)/\alpha},\nonumber \\
	& A_\gamma = A e^{(n_0-\gamma)/\alpha} \left( \gamma / n_0\right)^3.
\end{align}
The coefficients $g_n$ are given by the formulas:
\begin{align}
	& g_0 = \ln T_0, \qquad g_1 = T_1/T_0, \qquad g_2 = T_2/T_0 - g_1^2,  \nonumber \\
	& g_3 = T_3/T_0 - g_1^3 - 3g_1 g_2, \quad \qquad  g_4 = T_4/T_0 - g_1^4 - 6 g_1^2 g_2 - 4 g_1 g_3 - 3 g_2^2,
\end{align}
where $T_n(p,\alpha^*)$ are the following special functions
\begin{equation}
	T_n(p,\alpha^*) = \sum_{m=0}^{\infty} \frac{(\alpha^*)^m}{m!} m^n e^{pm^2}.
\end{equation}
Here $\alpha^*=v e^{\beta_c\mu^*}$, and the parameter $p$ has the form
\begin{equation}
	p = \frac{\beta_c}{2} \Phi^{(r)}(0) [\chi_0 + A_\gamma].
\end{equation} 
Note that we denote $\beta_c$ as the inverse critical temperature. Since $p$ fails to be the function of temperature, the coefficients $g_n$ are also temperature-free. The quantity $\tilde D(0)$ which has already appeared in both equations \eqref{eq:eos} and \eqref{3d2} is a function of temperature
\begin{equation}
	\tilde D(0) = g_2 - \frac{1}{2} \frac{g_3^2}{g_4} - \frac{1}{\beta W(0)},
\end{equation}
which equals zero at $T=T_c$. This circumstance gives us an expression for determining the critical temperature 
\begin{equation}
	T_c = \left(g_2 - \frac{1}{2} \frac{g_3^2}{g_4} \right) (B - 1 + \chi_0) \Phi^{(r)}(0).
\end{equation}    
The function $E_\mu(M, T)$ from the equation \eqref{eq:eos} is provided by
\begin{align}
	& E_\mu (M, T) = - \frac{1}{2 N_v} \ln (2\pi \beta W(0)) +  g_0 - \frac{\beta W(0)}{2} \!\! \left( \! \frac{\tilde\mu}{W(0)} \! \right)^{\! 2} \!\!\! - \frac{g_3}{g_4} {M} \! - \frac{g_3^2}{2 g_4^2}  \tilde D(0) - \frac{1}{24} \frac{g_3^4}{g_4^3}. 
\end{align}
Note that for all $ \tau> 0 $ there are no restrictions on the value of $ \bar \rho_0 $, which is a function of the temperature and the chemical potential
\begin{equation}\label{2d18}
	\bar \rho_0 = \left(- \frac{3\bar{M}}{g_4} + \sqrt{Q_t}\right)^{1/3} - \left(  \frac{3\bar{M}}{g_4} + \sqrt{Q_t} \right)^{1/3} \!\!\!\! ,
\end{equation}
where
\begin{equation}
	Q_t = \left(  \frac{2\tilde D(0)}{g_4}\right)^3 + \left( -\frac{3\bar{M}}{g_4}\right)^2, \qquad g_4<0.
\end{equation}  
...
\begin{equation}\label{3d3}
	\bar n = n_g - \bar{M} + \tilde g_2 \gamma_\tau \bar \rho_0,
\end{equation}
where
\begin{equation}\label{3d4}
	n_g = g_1 + n_c \tilde g_2 - \frac{1}{6} g_3^3 / g_4^2.
\end{equation}
...

### 2024-10-22 main_widom_morse.tex
add description of Fig2a, Fig2b in text

@@
 This happens because near the critical point, the system experiences very large fluctuations in density, and these fluctuations occur over increasingly large distances.
-Although, the equation of state is derived in the mean-field-like approximation,   in~\cite{DobushEtAl2024} we established that the isothermal compressibility of our cell fluid model diverge in the critical point, and consequently it display maxima in the supercritical region (see Figures~\ref{fig3}).
+Although, the equation of state~\eqref{eq:eosPTM_reduced} is derived in the mean-field-like approximation,   in~\cite{DobushEtAl2024} we established that the isothermal compressibility of our cell fluid model diverge in the critical point, and consequently it display maxima in the supercritical region (see Figures~\ref{fig3}).
 It means the possibility to plot the Widom line, and to expand the phase diagram in Figure~\ref{fig2} beyond the critical point.
@@
 \end{eqnarray}
-
-**************************************************************************
-
-[Figures only. Spots of text in progress.]
-
 \begin{figure}[h!] 
@@
 \end{figure}
+Behavior of the reduced isothermal compressibility versus effective chemical potential at various fixed values of the reduced temperature is represented graphically in Figure~\ref{fig3}a (wide range of $M$), and Figure~\ref{fig3}b (narrow range around $M=0$).
+Right at the critical point the compressibility diverge.
+As a result, there is still a "memory" of the critical point, where $\kappa_T$ reach local maxima, but once the system moves further into the supercritical region, these maxima become less pronounced, and the compressibility decreases.
+Below the critical temperature, negative values of $M$ correspond to gaseous densities, while positive values of $M$ represent liquid densities.
+Above the critical point, the fluid becomes continuous.
+However, Figure~\ref{fig2}a clearly shows significantly higher isothermal compressibility in the region where $M<0$, near the base of the maxima ridge, compared to the compressibility values in the $M>0$ region.
+In the first case ($M<0$), the supercritical fluid exhibits gas-like behavior.
+In the second case ($M>0$), it behaves like a liquid.
+Far beyond the critical point ($\tau > 0.5$), there is no noticeable difference in $\kappa_T$ values around the effective chemical potential $M=0$.
+At $\tau = 0.2$ the compressibility maxima become almost undistinguished (see Figure~\ref{fig2}b), they completely vanish at $\tau > 0.28$.
+
+**************************************************************************
+
+[Figure only. Spots of text in progress.]
 

### 2024-10-23 main_widom_morse.tex
start adding description of Fig4 in text

@@
 Above the critical point, the fluid becomes continuous.
-However, Figure~\ref{fig2}a clearly shows significantly higher isothermal compressibility in the region where $M<0$, near the base of the maxima ridge, compared to the compressibility values in the $M>0$ region.
+However, Figure~\ref{fig3}a clearly shows significantly higher isothermal compressibility in the region where $M<0$, near the base of the maxima ridge, compared to the compressibility values in the $M>0$ region.
 In the first case ($M<0$), the supercritical fluid exhibits gas-like behavior.
@@
 Far beyond the critical point ($\tau > 0.5$), there is no noticeable difference in $\kappa_T$ values around the effective chemical potential $M=0$.
-At $\tau = 0.2$ the compressibility maxima become almost undistinguished (see Figure~\ref{fig2}b), they completely vanish at $\tau > 0.28$.
+At $\tau = 0.2$ the compressibility maxima become almost undistinguished (see Figure~\ref{fig3}b), they completely vanish at $\tau > 0.28$.
 
-**************************************************************************
+To identify the Widom line maximize the isothermal compressibility by solving the equation
+\begin{equation}\label{eq:maximize}
+	\frac{\partial \kappa_T^*}{\partial M} = 0.
+\end{equation}
+Applying the solution of \eqref{eq:maximize} in the equation of state \eqref{def:reduced_pres}, one locates the Widom line position in the pressure-temperature plane (see Figure~\ref{fig4}, dashed line), expanding the phase diagram displayed in Figure~\ref{fig2} beyond the critical point.
 
-[Figure only. Spots of text in progress.]
+*******************************************************************************
+
+[more text in progress]
 
@@
 	\centering \includegraphics[width=0.5\textwidth]{f4.pdf}
-	\vskip-3mm\caption{Complete phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse}.
-Comparison of the Widom line (maxima locus of isothermal compressibility) and the crossover line between gas-like and liquid-like structures of the supercritical cell fluid based on the equation of state~\eqref{eq:eosPTM_reduced}.
+	\vskip-3mm\caption{Complete pressure-temperature phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse}.
+Comparison of the Widom line (maxima locus of isothermal compressibility) and the crossover between gas-like and liquid-like structures of the supercritical cell fluid based on combination of the equation of state~\eqref{eq:eosPTM_reduced} and the condition $M = 0$.
 An inset in the right bottom corner focuses on a range of $\tau$ around its critical value $-0.01 < \tau < 0.05$.	}\label{fig4}

### 2024-10-27 main_widom_morse.tex

@@
 At the critical point, the distinction between the liquid and gas phases of a substance disappears, and beyond this point, the substance exists in a single continuous supercritical state.
-Despite the absence of first-order phase transitions above the critical point, fluids exhibit distinct liquid-like and gas-like structures under supercritical conditions, and display a unique mix of properties.
+Despite the absence of first-order phase transitions above the critical point, fluids exhibit distinct liquid-like and gas-like behavior under supercritical conditions, and display a unique mix of properties.
 With densities comparable to liquids they possess low viscosity and high diffusivity like gases resulting in high solvent power.
 Moreover, the supercritical properties are instantly tunable by changing pressure and temperature.
-This is the reason why supercritical fluids have diverse applications across various industries \cite{KnezEtAl2014}.
+This is the reason why supercritical fluids have diverse applications across various industries \cite{KMLetal14}.
 
@@
 All three concepts provide a way to describe how fluids behave in the supercritical region, where the traditional liquid-gas phase distinction no longer applies.
-The Frenkel line \cite{BrykEtAl2017} focuses on the dynamics of molecular motion, separating regimes where the fluid behaves more like a liquid (with oscillatory molecular motion) from those where it behaves more like a gas (with only diffusive motion).
-The percolation line \cite{CampiEtAl2001} is related to spatial structures, marking the threshold where clusters of fluid or density fluctuations start forming a continuous, system-spanning network.
+The Frenkel line \cite{BGMetal17} focuses on the dynamics of molecular motion, separating regimes where the fluid behaves more like a liquid (with oscillatory molecular motion) from those where it behaves more like a gas (with only diffusive motion).
+The percolation line \cite{CKS01} is related to spatial structures, marking the threshold where clusters of fluid or density fluctuations start forming a continuous, system-spanning network.
 
-In the present work we are particularly interested in the Widom line \cite{XuEtAl2005,McMillanStanley2010,SimeoniEtAl2010} which focuses on thermodynamic properties and represents a conditional boundary between liquid-like and gas-like regions in the supercritical fluid, based on response functions.
-Thus the Widom line \--- the locus of correlation length maxima emanating from the critical point \cite{XuEtAl2005}, is treated as the continuation of the coexistence line above the critical point.
+In the present work we are particularly interested in the Widom line \cite{XKBetal05,MS10,AKM17} which focuses on thermodynamic properties and represents a conditional boundary between liquid-like and gas-like regions in the supercritical fluid, based on response functions.
+Thus the Widom line \--- the locus of correlation length maxima emanating from the critical point \cite{XKBetal05}, is treated as the continuation of the coexistence line above the critical point.
 Some thermodynamic response functions, for example, isobaric heat capacity and isothermal compressibility, also exhibit maxima in the supercritical region.
-In the critical region beyond the critical point the maxima lines of the correlation length and response functions virtually merge into a single line \cite{ArtemenkoEtAl2017} and then diverge from each other as the temperature rise.
-Thus in the critical region, the Widom line is viewed as the maxima locus of thermodynamic response functions \cite{Proctor2020Book}.
+In the critical region beyond the critical point the maxima lines of the correlation length and response functions virtually merge into a single line \cite{AKM17} and then diverge from each other as the temperature rise.
+Thus in the critical region, the Widom line is viewed as the maxima locus of thermodynamic response functions \cite{Proctor20}.
 
@@
 % we would like to share an idea
+
+[Here a short paragraph that this work is a logical consequence of our previous works on the equation of state and response functions for the cell fluid model with interaction expressed by the modified Morse potential]
 
@@
 
-In the present work we take advantage of the results obtained in \cite{KozlovskiiDobush2020,DobushEtAl2024}.
-In \cite{KozlovskiiDobush2020}, the phase space of collective variables%\footnote{The term "collective variables" is commonly used to refer to a specific category of variables unique to each physical system.
+In the present work we take advantage of the results obtained in \cite{KD20,DKRP24arxiv,DKRP24}.
+In \cite{KD20}, the phase space of collective variables%\footnote{The term "collective variables" is commonly used to refer to a specific category of variables unique to each physical system.
 In the context of liquid-vapour systems, these collective variables are linked to modes of density deviation from the critical density.
@@
 The resulting EoSs are applicable in a wide range of temperatures below and above the critical point.
-In \cite{DobushEtAl2024}, we used these equations as the basis for the calculation and graphical representation of the thermodynamic response functions (isothermal compressibility, thermal pressure coefficient, thermal expansion) of the cell fluid model with modified Morse potential in the supercritical region.
+In \cite{DKRP24arxiv,DKRP24}, we used these equations as the basis for the calculation and graphical representation of the thermodynamic response functions (isothermal compressibility, thermal pressure coefficient, thermal expansion) of the cell fluid model with modified Morse potential in the supercritical region.
 
@@
 where $\varepsilon$ is the depth of the potential well at $r=R_0$.
-Applying $\gamma=2$ reduce the  to the ordinary Morse potential~\cite{Morse1929} at $\gamma=2$.
-For a more detailed discussion of such modified Morse potential, see Sections~1 and~2 in~\cite{KozlovskiiDobush2020}.
+Applying $\gamma=2$ reduces $U(r)$ to the ordinary Morse potential~\cite{Morse29}.
+For a more detailed discussion of such modified Morse potential, see Sections~1 and~2 in~\cite{KD20}.
 
-Our modification contains an additional repulsive term, which enables us to single out a reference system (in the reciprocal space) and apply the method of collective variables to calculating the grand partition function~\cite{KozlovskiiDobush2016}.
+Our modification contains an additional repulsive term, which enables us to single out a reference system (in the reciprocal space) and apply the method of collective variables to calculating the grand partition function~\cite{KD16}.
 
@@
 
-The equation of state obtained in~\cite{KozlovskiiDobush2020} reads
+The equation of state obtained in~\cite{KD20} reads
 \begin{equation}
@@
 
-The parameter $\chi_0$ is used in~\cite{KozlovskiiDobush2020} to single out a contribution in the Fourier transform of the potential that is treated as a reference system defined in the reciprocal space, and is selected as $\chi_0 = 0.07$~\cite{KozlovskiiDobush2020}.
+The parameter $\chi_0$ is used in~\cite{KD20} to single out a contribution in the Fourier transform of the potential that is treated as a reference system defined in the reciprocal space, and is selected as $\chi_0 = 0.07$~\cite{KD20}.
 
@@
 
-The quantity $\beta_c$ denotes the critical value of the inverse temperature. In~\cite{KozlovskiiDobush2020} it was found that
+The quantity $\beta_c$ denotes the critical value of the inverse temperature. In~\cite{KD20} it was found that
 \begin{equation*}
@@
 Since $p$ is independent of temperature, the coefficients $g_n$ are also independent of temperature.
-The numerical values for other coefficients used in this paper are the same as those in~\cite[eqs.~(5), (23), and~(24)]{KozlovskiiDobush2020}:
+The numerical values for other coefficients used in this paper are the same as those in~\cite[eqs.~(5), (23), and~(24)]{KD20}:
 \begin{eqnarray}
@@
 \end{equation}
-The condition $d = 0$ defines the critical temperature~\cite{KozlovskiiDobush2020}
+The condition $d = 0$ defines the critical temperature~\cite{KD20}
 \begin{equation}
@@
 	\includegraphics[width=0.45\textwidth]{f1.pdf} 
-	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the effective chemical potential $M$ at $T \geq T_c$ represented by black lines.
-Grey lines correspond to isotherms of pressure at $T < T_c$ based on the results for the same model taken from~\cite{KozlovskiiDobush2020}.
+	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the effective chemical potential $M$. Black lines correspond to $T \geq T_c$. Grey lines correspond to $T < T_c$. 
 	}\label{fig1}
@@
 Figure~\ref{fig1} illustrates the relationship between the reduced pressure $P^*$ and the effective chemical potential $M$ for various relative temperature values $\tau$.
-Below the critical temperature ($\tau < 0$), іsotherms are curves consisting of two segments connected by a vertex (grey lines in Figure~\ref{fig1}).
-This form indicates a first-order phase transition occurs in the system.
-Within our approach~\cite{KozlovskiiDobush2020} at $T<T_c$, negative values of the effective chemical potential mean the system is in a gaseous state.
-The vertex $M=0$ is a point of phase coexistence.
+Below the critical temperature ($\tau < 0$), each іsotherm contains a kink (grey lines in Figure~\ref{fig1}).
+This form indicates a first-order phase transition in the system.
+Within our approach~\cite{KD20} at $T<T_c$, negative values of the effective chemical potential $M$ mean the system is in a gaseous state.
+The condition $M=0$ is a point of phase coexistence.
 Positive values of the effective chemical potential correspond to the liquid state.
 In both cases, the pressure tends to increase with $M$, but the slope of the liquid isotherm is much sharper.
-At the critical temperature pressure is a smooth plane curve (black dotted line in Figure~\ref{fig1}) with an inflection point at $M=0$.
+At the critical temperature, the pressure is a smooth curve (black dotted line in Figure~\ref{fig1}) with an inflection point at $M=0$.
 Thus, at the critical point $M=0, \tau=0$, $P^*_c = 1.606$.
 Beyond the critical temperature, the pressure behaves as a monotonically increasing function, and no phase transitions are observed.
-Summing up, since below the critical temperature, $M=0$ stands for phase coexistence, we obtain a phase diagram in pressure-temperature terms (see Figure~\ref{fig2}) by simply using this condition in the equation of state \eqref{eq:eosPTM_reduced}.
+Summing up, since below the critical temperature, $M=0$ is the condition for phase coexistence, we obtain a phase diagram in pressure-temperature terms (see Figure~\ref{fig2}) by simply using this condition in the equation of state \eqref{eq:eosPTM_reduced}.
 \begin{figure}[t!]
@@
 	\includegraphics[width=0.45\textwidth]{f2.pdf} 
-	\vskip-3mm\caption{Phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse} based on the equation of state~\eqref{eq:eosPTM_reduced}. 
+	\vskip-3mm\caption{Phase diagram in the $P^*-\tau$ plane for the cell fluid model with the interaction potential \eqref{def:mod_morse}.
+The result is based on the equation of state~\eqref{eq:eosPTM_reduced}.
 	}\label{fig2}
@@
 Instead, it exists as a single, continuous phase with no sharp liquid-gas boundary.
-Nevertheless, fluids exhibit distinct liquid-like and gas-like structures beyond the critical temperature.
-A way to represent a conditional boundary between liquid-like and gas-like regions in the supercritical fluid is the Widom line \--- a continuation of the coexistence line based on the maxima locus of some thermodynamic response functions~\cite{XuEtAl2005,McMillanStanley2010,SimeoniEtAl2010}.
+Nevertheless, fluids exhibit distinct liquid-like and gas-like behavior beyond the critical temperature.
+A way to represent a conditional boundary between liquid-like and gas-like regions in the supercritical fluid is the Widom line \--- a continuation of the coexistence line based on the maxima locus of some thermodynamic response functions~\cite{XKBetal05,MS10,SBGetal10}.
 This happens because near the critical point, the system experiences very large fluctuations in density, and these fluctuations occur over increasingly large distances.
-Although, the equation of state~\eqref{eq:eosPTM_reduced} is derived in the mean-field-like approximation,  in~\cite{DobushEtAl2024} we established that the isothermal compressibility of our cell fluid model diverge in the critical point, and consequently it display maxima in the supercritical region (see Figures~\ref{fig3}).
-It means the possibility to plot the Widom line, and to expand the phase diagram in Figure~\ref{fig2} beyond the critical point.
+Although, the equation of state~\eqref{eq:eosPTM_reduced} is derived in the mean-field-like approximation,  in~\cite{DKRP24arxiv,DKRP24} we established that the isothermal compressibility of the cell fluid model diverge in the critical point, and consequently it displays maxima in the supercritical region (see Figures~\ref{fig3}).
+It means the possibility to derive the Widom line, and to expand the phase diagram in Figure~\ref{fig2} beyond the critical point.
  In what follows we consider this in more detail.
@@
 \end{equation}
-In \cite{DobushEtAl2024} we derived expression for the reduced isothermal compressibility $\kappa^*_T \equiv \dfrac{\varepsilon \kappa_T}{v}$  in terms of thermodynamic derivatives suitable for the reduced form of equation of state in pressure-temperture-chemical potential terms \eqref{eq:eosPTM_reduced}
+In \cite{DKRP24arxiv,DKRP24} we derived expression for the reduced isothermal compressibility $\kappa^*_T \equiv \dfrac{\varepsilon \kappa_T}{v}$  in terms of thermodynamic derivatives suitable for the reduced form of the equation of state in pressure-temperture-chemical potential terms \eqref{eq:eosPTM_reduced}
 It is expressed in terms of the reduced pressure $P^*$ and the reduced particle number density $\rho^*$ as follows
@@
 \end{equation}
-The quantity $\rho^*_c$ in the equation \eqref{eq:density} is the critical density~\cite{KozlovskiiDobush2020,DobushEtAl2024}
+The quantity $\rho^*_c$ in the equation \eqref{eq:density} is the critical density~\cite{KD20,DKRP24arxiv,DKRP24}
 \begin{eqnarray}\label{eq:crit_dens}
@@
 Behavior of the reduced isothermal compressibility versus effective chemical potential at various fixed values of the reduced temperature is represented graphically in Figure~\ref{fig3}a (wide range of $M$), and Figure~\ref{fig3}b (narrow range around $M=0$).
-Right at the critical point the compressibility diverge.
-As a result, there is still a "memory" of the critical point, where $\kappa_T$ reach local maxima, but once the system moves further into the supercritical region, these maxima become less pronounced, and the compressibility decreases.
+Right at the critical point the compressibility diverges.
+As a result, there is still a "memory" of the critical point, where $\kappa_T$ reaches local maxima.
+However, once the system moves further into the supercritical region, these maxima become less pronounced, and the compressibility decreases.
 Below the critical temperature, negative values of $M$ correspond to gaseous densities, while positive values of $M$ represent liquid densities.
@@
 
-To identify the Widom line maximize the isothermal compressibility by solving the equation
+To identify the Widom line, the locus of the isothermal compressibility maxima is found by solving the equation
 \begin{equation}\label{eq:maximize}
-	\frac{\partial \kappa_T^*}{\partial M} = 0.
+	\frac{\partial \kappa_T^*}{\partial M} = 0
 \end{equation}
-Applying the solution of \eqref{eq:maximize} in the equation of state \eqref{def:reduced_pres}, one locates the Widom line position in the pressure-temperature plane (see Figure~\ref{fig4}, dashed line), expanding the phase diagram displayed in Figure~\ref{fig2} beyond the critical point.
+with respect to $M$. 
+Substituting the solution of \eqref{eq:maximize} into the equation of state \eqref{def:reduced_pres}, one locates the Widom line in the pressure-temperature plane (see Figure~\ref{fig4}, dashed line), expanding the phase diagram displayed in Figure~\ref{fig2} beyond the critical point.
 
@@
 \vskip3mm \textit{This work was supported by the National Research Foundation of Ukraine under the project No.~2023.03/0201.}
+
+%\bibliographystyle{elsarticle-num}
+\bibliographystyle{plain}
+\bibliography{extracted}
+
+\end{document}
 
@@
 
-\end{document}
+%\end{document}
 

### 2024-10-28 main_widom_morse.tex
add 2 sentences on M=0 in the supercrit region
removed f3a figure, changed the text respectively, removed b) from figure f3b

@@
 \begin{figure}[h!] 
-	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
-	\includegraphics[width=0.46\textwidth]{f3bb.pdf}
+%	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
+	\centering \includegraphics[width=0.46\textwidth]{f3bb.pdf}
 	%\includegraphics[width=\column]{f2b.pdf}
@@
 	\caption{The reduced isothermal compressibility $\kappa^*_T$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$.
-The two figures differ in the scale of $M$.
-Part~(a) covers a wider range of $M$.
-Part~(b) focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\kappa^*_T$.
+The plot focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\kappa^*_T$.
 	}
@@
 \end{figure}
-Behavior of the reduced isothermal compressibility versus effective chemical potential at various fixed values of the reduced temperature is represented graphically in Figure~\ref{fig3}a (wide range of $M$), and Figure~\ref{fig3}b (narrow range around $M=0$).
+Behavior of the reduced isothermal compressibility versus effective chemical potential at various fixed values of the reduced temperature is represented graphically in Figure~\ref{fig3} (narrow range around $M=0$).
 Right at the critical point the compressibility diverges.
@@
 Above the critical point, the fluid becomes continuous.
-However, Figure~\ref{fig3}a clearly shows significantly higher isothermal compressibility in the region where $M<0$, near the base of the maxima ridge, compared to the compressibility values in the $M>0$ region.
+However, Figure~\ref{fig3} clearly shows significantly higher isothermal compressibility in the region where $M<0$, near the base of the maxima ridge, compared to the compressibility values in the $M>0$ region.
 In the first case ($M<0$), the supercritical fluid exhibits gas-like behavior.
@@
 Far beyond the critical point ($\tau > 0.5$), there is no noticeable difference in $\kappa_T$ values around the effective chemical potential $M=0$.
-At $\tau = 0.2$ the compressibility maxima become almost undistinguished (see Figure~\ref{fig3}b), they completely vanish at $\tau > 0.28$.
+At $\tau = 0.2$ the compressibility maxima become almost undistinguished, they completely vanish at $\tau > 0.28$.
 
@@
 
-*******************************************************************************
-
-[more text in progress]
+On the other hand, within the present approach, $M=0$ at the points of phase coexistance up to the critical point.
+Thus, the implementation of the condition  $M=0$ appears to be a natural continuation of the coexistance line into the supercritical region.
 

### 2024-10-28 main_widom_morse.tex
work on introduction

@@
 
-% directly from the equation of state 
-[Here will be a small paragraph about challenges on developing the global equation of state directly predicting crossover line in supercritical fluids]
-% we would like to share an idea
+In our previous work~\cite{KD20}, the equation of state for a many-particle system interacting via a modified Morse potential was developed in the framework of a cell model.
+Recently~\cite{DKRP24,DKRP24arxiv}, the thermodynamic response functions, namely the isothermal compressibility, the thermal pressure coefficient, and the thermal expansion coefficient, have been calculated for the same model system.
+The isothermal compressibility and thermal expansion coefficient both develop a maximum in some region beyond critical point.
+On the one hand, such response function maxima, as mentioned above, are considered as good candidates for crossover lines between gas-like and liquid-like behaviors in the supercritical region.
+On the other hand, they are also considered as natural continuation of the coexistence line.
+Since our model allows for analytical expressions for many thermodynamic quantities~\cite{KD20,DKRP24}, it is straightforward to calculate and build such continuations.
+As a logical sequence of our previous works for the cell model with the modified Morse potential, in this paper we present results for two lines that can be cosidered as continuations of the coexistence line beyond the critical point in the pressure-temperature plane.
+The first one is the locus of the isothermal compressibility maxima.
+This line, as will be shown, ends (disappears, ceases to exist) at some temperature above the critical one.
+The second one is a line that is derived from the same condition that the first-order phase transition line below the critical point - the condition on the effective chemical potential $M=0$.
+This line can be extended into the supercritical region without any limits in temperature or pressure.
 
-[Here a short paragraph that this work is a logical consequence of our previous works on the equation of state and response functions for the cell fluid model with interaction expressed by the modified Morse potential]
-
-In Section~\ref{sec:eos}, we present the modified Morse potential used in our work, the explicit expression for the equation of state, and plot of pressure isotherms at temperatures below and above the critical one.
-In Section~\ref{sec:widom}, we display the Widom line as the maxima locus of the isothermal compressibility and introduce the supercritical crossover line obtained directly from the equation of state of the cell fluid model.
+In Section~\ref{sec:eos}, we recall the form of the modified Morse potential, present the explicit expression for the equation of state, and illustrate pressure isotherms at temperatures below and above the critical one.
+In Section~\ref{sec:widom}, we display the Widom line as the maxima locus of the isothermal compressibility and introduce the supercritical crossover line obtained directly from the equation of state of the cell fluid model applying the proper condition on the effective chemical potential.
 
@@
 The resulting EoSs are applicable in a wide range of temperatures below and above the critical point.
-In \cite{DKRP24arxiv,DKRP24}, we used these equations as the basis for the calculation and graphical representation of the thermodynamic response functions (isothermal compressibility, thermal pressure coefficient, thermal expansion) of the cell fluid model with modified Morse potential in the supercritical region.
+In~\cite{DKRP24arxiv,DKRP24}, we used these equations as the basis for the calculation and graphical representation of the thermodynamic response functions (isothermal compressibility, thermal pressure coefficient, thermal expansion) of the cell fluid model with modified Morse potential in the supercritical region.
 
@@
 %\bibliographystyle{elsarticle-num}
-\bibliographystyle{plain}
+%\bibliographystyle{plain}
+\bibliographystyle{gost2008}
 \bibliography{extracted}

### 2024-10-29 main_widom_morse.tex
refactor intro

@@
 \section{Introduction}
-A supercritical fluid is a state of matter that occurs when a substance is subjected to temperatures and pressures above its critical point.
-Below the critical point the substance exhibits a liquid-gas coexistence.
-At the critical point, the distinction between the liquid and gas phases of a substance disappears, and beyond this point, the substance exists in a single continuous supercritical state.
+A supercritical fluid is a state of matter that occurs when a substance is subjected to temperatures and pressures above its critical point (CP).
+Below CP the substance exhibits a liquid-gas coexistence.
+At CP, the distinction between the liquid and gas phases of a substance disappears, and beyond this point, the substance exists in a single continuous supercritical state.
 Despite the absence of first-order phase transitions above the critical point, fluids exhibit distinct liquid-like and gas-like behavior under supercritical conditions, and display a unique mix of properties.
@@
 
-In the context of supercritical fluids, the Widom line, Frenkel line, and percolation line are theoretical and physical constructs used to describe the complex behavior of fluids beyond the critical point.
-All three concepts provide a way to describe how fluids behave in the supercritical region, where the traditional liquid-gas phase distinction no longer applies.
-The Frenkel line \cite{BGMetal17} focuses on the dynamics of molecular motion, separating regimes where the fluid behaves more like a liquid (with oscillatory molecular motion) from those where it behaves more like a gas (with only diffusive motion).
-The percolation line \cite{CKS01} is related to spatial structures, marking the threshold where clusters of fluid or density fluctuations start forming a continuous, system-spanning network.
+Although the traditional liquid-gas phase distinction does not apply in the supercritical region, scientists continue to search for crossover regions or definitive lines that distinguish liquid-like and gas-like behaviors in the phase diagram.
+Naturally, such boundaries originate from CP and may either vanish away from it or extend infinitely far.
+Among known approaches to this problem are the Widom line, which represents the locus of correlation length maxima~\cite{FW69}; the Frenkel line~\cite{BGMetal17}, focusing on the dynamics of molecular motion; and the percolation line~\cite{CKS01}, related to spatial structures and marking the threshold where clusters of fluid or density fluctuations form a continuous, system-spanning network.
 
-In the present work we are particularly interested in the Widom line \cite{XKBetal05,MS10,AKM17} which focuses on thermodynamic properties and represents a conditional boundary between liquid-like and gas-like regions in the supercritical fluid, based on response functions.
-Thus the Widom line \--- the locus of correlation length maxima emanating from the critical point \cite{XKBetal05}, is treated as the continuation of the coexistence line above the critical point.
-Some thermodynamic response functions, for example, isobaric heat capacity and isothermal compressibility, also exhibit maxima in the supercritical region.
-In the critical region beyond the critical point the maxima lines of the correlation length and response functions virtually merge into a single line \cite{AKM17} and then diverge from each other as the temperature rise.
-Thus in the critical region, the Widom line is viewed as the maxima locus of thermodynamic response functions \cite{Proctor20}.
+In the present work, we focus on the Widom line \cite{XKBetal05,MS10,AKM17}, a concept based on thermodynamic response functions that represents a conditional boundary between liquid-like and gas-like regions in supercritical fluids.
+Thus, the Widom line \--- the locus of correlation length maxima emanating from CP~\cite{XKBetal05} \--- is treated as the continuation of the coexistence line above CP.
+Moreover, some thermodynamic response functions, such as isobaric heat capacity and isothermal compressibility, also exhibit maxima in the supercritical region.
+In the critical region beyond the critical point, the maxima lines of the correlation length and these response functions virtually merge into a single line~\cite{AKM17} and then diverge from each other as the temperature increases.
+Therefore, in the critical region, the Widom line is viewed as the locus of maxima of thermodynamic response functions~\cite{Proctor20}.
 
@@
 
-\section{\label{sec:eos}The equation of state: pressure as a function of temperature and chemical potential}
+\section{\label{sec:eos} The equation of state}
 
-In the present work we take advantage of the results obtained in \cite{KD20,DKRP24arxiv,DKRP24}.
-In \cite{KD20}, the phase space of collective variables%\footnote{The term "collective variables" is commonly used to refer to a specific category of variables unique to each physical system.
-In the context of liquid-vapour systems, these collective variables are linked to modes of density deviation from the critical density.
-This set of collective variables encompasses those which average values are connected to order parameters.Thus, the depiction of the phase transition within the phase space of collective variables appears compelling.}
+In the present work we take advantage of the results obtained in \cite{KD20,DKRP24arxiv,DKRP24}. In \cite{KD20}, the phase space of collective variables
 is used for calculating the grand partition function of a cell fluid model in the zero-mode approximation and consequently deriving the explicit forms of the equation of state (EoS) both in pressure-temperature-chemical potential terms and pressure-temperature-density terms.
@@
 
-\section{\label{sec:widom}Crossover line between gas-like and liquid-like structures of the supercritical fluid. The Widom line}
+\section{\label{sec:widom}Crossover lines between gas-like and liquid-like behavior of the supercritical fluid.}
 

### 2024-10-29 main_widom_morse.tex
add abstract and conclusion

@@
 \subsection*{Abstract}
+A cell fluid model with a modified Morse potential is studied.
+The supercritical states are considered with respect to a possibility to build a separation boundary between liquid-like and gas-like bahaviors.
+Two different lines are calculated that can be used for this purpose: the locus of the isothermal compressibility maxima and the line where the effective chemical potential is zero, $M=0$.
+The line corresponding to the condition $M=0$ is analoguous to the Ising model in zero external field.
 
@@
 
-In the present work we take advantage of the results obtained in \cite{KD20,DKRP24arxiv,DKRP24}. In \cite{KD20}, the phase space of collective variables
-is used for calculating the grand partition function of a cell fluid model in the zero-mode approximation and consequently deriving the explicit forms of the equation of state (EoS) both in pressure-temperature-chemical potential terms and pressure-temperature-density terms.
-The resulting EoSs are applicable in a wide range of temperatures below and above the critical point.
-In~\cite{DKRP24arxiv,DKRP24}, we used these equations as the basis for the calculation and graphical representation of the thermodynamic response functions (isothermal compressibility, thermal pressure coefficient, thermal expansion) of the cell fluid model with modified Morse potential in the supercritical region.
+In the present work, we utilize the results obtained in~\cite{KD20,DKRP24arxiv,DKRP24}.
+In~\cite{KD20}, the phase space of collective variables was used to calculate the grand partition function of a cell fluid model in the zero-mode approximation, leading to explicit forms of the equation of state (EoS) in both pressure-temperature-chemical potential and pressure-temperature-density terms.
+These EoSs are applicable across a wide range of temperatures below and above the CP.
+Furthermore, in~\cite{DKRP24arxiv,DKRP24}, we have used these equations as the basis for calculating and graphically representing the thermodynamic response functions \--- specifically, the isothermal compressibility, thermal pressure coefficient, and thermal expansion \--- of the cell fluid model with the modified Morse potential in the supercritical region.
 
@@
 
-The potential of interaction between particles is taken in a form of a modified Morse potential
+The modified Morse potential used to describe the interaction between particles is given by 
 \begin{eqnarray}
@@
 Applying $\gamma=2$ reduces $U(r)$ to the ordinary Morse potential~\cite{Morse29}.
-For a more detailed discussion of such modified Morse potential, see Sections~1 and~2 in~\cite{KD20}.
-
-Our modification contains an additional repulsive term, which enables us to single out a reference system (in the reciprocal space) and apply the method of collective variables to calculating the grand partition function~\cite{KD16}.
+For a more detailed discussion of such modified Morse potential, see Sections~1 and~2 in~\cite{KD20}, and Section~2 in~\cite{DKRP24arxiv}.
 
@@
 \end{equation}  
-
 The function $E_\mu(M, T)$ from the equation \eqref{eq:eosMT} is provided by
@@
 Figure~\ref{fig1} illustrates the relationship between the reduced pressure $P^*$ and the effective chemical potential $M$ for various relative temperature values $\tau$.
-Below the critical temperature ($\tau < 0$), each іsotherm contains a kink (grey lines in Figure~\ref{fig1}).
+Below the critical temperature ($\tau < 0$), each іsotherm contains a kink (three bottom lines in Figure~\ref{fig1}).
 This form indicates a first-order phase transition in the system.
-Within our approach~\cite{KD20} at $T<T_c$, negative values of the effective chemical potential $M$ mean the system is in a gaseous state.
+Within our approach~\cite{KD20} at $T<T_c$, negative values of the effective chemical potential $M$ imply the system is in a gaseous state.
 The condition $M=0$ is a point of phase coexistence.
 Positive values of the effective chemical potential correspond to the liquid state.
-In both cases, the pressure tends to increase with $M$, but the slope of the liquid isotherm is much sharper.
+In both cases, the pressure tends to increase with $M$, but the slope of the liquid isotherm is much steeper.
 At the critical temperature, the pressure is a smooth curve (black dotted line in Figure~\ref{fig1}) with an inflection point at $M=0$.
-Thus, at the critical point $M=0, \tau=0$, $P^*_c = 1.606$.
+Thus, at the critical point coordinates are defined by $M=0, \tau=0$, $P^*_c = 1.606$.
 Beyond the critical temperature, the pressure behaves as a monotonically increasing function, and no phase transitions are observed.
-Summing up, since below the critical temperature, $M=0$ is the condition for phase coexistence, we obtain a phase diagram in pressure-temperature terms (see Figure~\ref{fig2}) by simply using this condition in the equation of state \eqref{eq:eosPTM_reduced}.
+Summing up, below the critical temperature, $M=0$ is the condition for phase coexistence.
+Substituring this condition into the EoS~\eqref{eq:eosPTM_reduced}, we obtain a phase diagram in pressure-temperature plane (see Figure~\ref{fig2}).
+We will use the same condition to obtain a continuation of the coexistence line into the supercritical region.
 \begin{figure}[t!]
@@
 
-%\section{Ентропія}
+\section{\label{sec:widom} Crossover lines between gas-like and liquid-like behavior of the supercritical fluid.}
 
-\section{\label{sec:widom}Crossover lines between gas-like and liquid-like behavior of the supercritical fluid.}
-
- In the supercritical region, the substance no longer behaves distinctly as either a liquid or a gas.
-Instead, it exists as a single, continuous phase with no sharp liquid-gas boundary.
-Nevertheless, fluids exhibit distinct liquid-like and gas-like behavior beyond the critical temperature.
-A way to represent a conditional boundary between liquid-like and gas-like regions in the supercritical fluid is the Widom line \--- a continuation of the coexistence line based on the maxima locus of some thermodynamic response functions~\cite{XKBetal05,MS10,SBGetal10}.
-This happens because near the critical point, the system experiences very large fluctuations in density, and these fluctuations occur over increasingly large distances.
-Although, the equation of state~\eqref{eq:eosPTM_reduced} is derived in the mean-field-like approximation,  in~\cite{DKRP24arxiv,DKRP24} we established that the isothermal compressibility of the cell fluid model diverge in the critical point, and consequently it displays maxima in the supercritical region (see Figures~\ref{fig3}).
-It means the possibility to derive the Widom line, and to expand the phase diagram in Figure~\ref{fig2} beyond the critical point.
+ In the supercritical region, substances no longer behave distinctly as either liquids or gases but exist as a single, continuous phase without a sharp liquid-gas boundary.
+However, fluids can still exhibit liquid-like and gas-like behaviors beyond the critical temperature.
+A way to represent a conditional boundary between these regions in supercritical fluids is the Widom line \--- a continuation of the coexistence line based on the locus of maxima of certain thermodynamic response functions~\cite{XKBetal05,MS10,SBGetal10}.
+This occurs because near CP, the system experiences significant fluctuations in density over increasingly large spatial scales.
+Although, the equation of state~\eqref{eq:eosPTM_reduced} is derived in the mean-field-like approximation, we established in~\cite{DKRP24arxiv,DKRP24} that the isothermal compressibility of the cell fluid model diverges at CP and consequently displays maxima in the supercritical region (see Figures~\ref{fig3}).
+This suggests the possibility of deriving the Widom line and expanding the phase diagram in Figure~\ref{fig2} beyond CP.
  In what follows we consider this in more detail.
@@
 \section{Conclusions}
-
-[Spots of text in progress.]
+The locus of the isothermal compressibility maxima is calculated for the cell fluid model with the modified Morse potential.
+This is considered as a good candidate for a separation line between gas-like and liquid-like regions in the supercritical phase.
+We also suggest another such line derived from the condition for the effective chemical potential $M=0$.
+This is a natural extension of the coexistence line below CP, since the same condition is used in both cases.
+A noticeable difference between the two separation lines is that the compressibility maxima end at some temperature (or pressure) value above CP, see Fig.~\ref{fig4}, while the second line extends infinitely far from CP.
+Investigation of how physical quantities change their behavior when crossing these lines will be the subject of next studies.
 

### 2024-10-29 main_widom_morse.tex

@@
 	\renewcommand{\theequation}{\arabic{section}}
-	\section*{Crossover line in the supercritical cell fluid model (I haven't thought up a good title yet)}
+	\section*{Supercritical Crossover Lines in the Cell Fluid Model}
 	\setcounter{equation}{0}
@@
 
-On the other hand, within the present approach, $M=0$ at the points of phase coexistance up to the critical point.
-Thus, the implementation of the condition  $M=0$ appears to be a natural continuation of the coexistance line into the supercritical region.
+On the one hand, within the present approach, the condition $M=0$ holds along phase coexistence up to the critical point.
+Thus, the implementation of the condition  $M=0$ above CP seems to be a natural continuation of the coexistence line into the supercritical region.
+On the other hand, taking into account the functional form of the grand partition function~\cite{KD16,KR09}, the condition of $M=0$ is equivalent to the absence of an external field in the Ising model.
+This analogy is interesting from the perspective of mapping between the Ising model and fluids~\cite{Kulinskii10jcp,BK11}.
 

### 2024-10-29 main_widom_morse.tex
added new figures in text

@@
 At $\tau = 0.2$ the compressibility maxima become almost undistinguished, they completely vanish at $\tau > 0.28$.
-
+\begin{figure}[h!] 
+	%	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
+	\centering \includegraphics[width=0.46\textwidth]{f3cc.pdf}
+	%\includegraphics[width=\column]{f2b.pdf}
+	%\includegraphics[width=\column]{f2c.pdf}
+	\vskip-3mm
+	\caption{The reduced thermal expansion $\alpha^*_P$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$.
+The plot focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\alpha^*_T$.
+	}
+	\label{fig3b}
+\end{figure}
 To identify the Widom line, the locus of the isothermal compressibility maxima is found by solving the equation
@@
 \begin{figure}[h!]
-	\centering \includegraphics[width=0.5\textwidth]{f4.pdf}
+	\centering \includegraphics[width=0.5\textwidth]{f4b.pdf}
 	\vskip-3mm\caption{Complete pressure-temperature phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse}.
-Comparison of the Widom line (maxima locus of isothermal compressibility) and the crossover between gas-like and liquid-like structures of the supercritical cell fluid based on combination of the equation of state~\eqref{eq:eosPTM_reduced} and the condition $M = 0$.
+Comparison of the Widom line (maxima locus of isothermal compressibility and thermal pressure) and the crossover between gas-like and liquid-like structures of the supercritical cell fluid based on combination of the equation of state~\eqref{eq:eosPTM_reduced} and the condition $M = 0$.
 An inset in the right bottom corner focuses on a range of $\tau$ around its critical value $-0.01 < \tau < 0.05$.	}\label{fig4}

### 2024-10-29 main_widom_morse.tex
add thermal expansion coefficient

@@
 The supercritical states are considered with respect to a possibility to build a separation boundary between liquid-like and gas-like bahaviors.
-Two different lines are calculated that can be used for this purpose: the locus of the isothermal compressibility maxima and the line where the effective chemical potential is zero, $M=0$.
-The line corresponding to the condition $M=0$ is analoguous to the Ising model in zero external field.
+Three different lines are calculated that can be used for this purpose: the locus of the isothermal compressibility maxima, the locus of the thermal expansion coefficient maxima, and the line where the effective chemical potential is zero, $M=0$.
+By the symmetry of the functionals for the partition functions, the condition $M=0$ in fluids is analogous to the absence of an external field in the Ising model.
 
@@
 
-**************************************************************
-
-[The following is the literal copy (compilation) from our recent preprint on response functions. Perhaps we should rephrase it, or change it somehow. Please help.]
-
-**************************************************************
-
 The modified Morse potential used to describe the interaction between particles is given by 
@@
 
-%The explicit form of the equation of state in terms of chemical potential-temperature in the case of $T\geq T_c$ is as follows:
+We use $\varepsilon$ and $v$ as natural units for energy and volume.
+Therefore, the following reduced quantities are defined: the reduced density $\rho^*=\rho v$, where $\rho = \langle N \rangle /V$; reduced temperature $T^* = k_{\rm B}T/\varepsilon$; reduced pressure $P^* = Pv/\varepsilon$; reduced isothermal compressibility $\kappa^*_{T} = \varepsilon \kappa_T/v$; and the reduced thermal expansion coefficient $\alpha^*_{P} = \varepsilon \alpha_{P} / k_{\rm B}$.
+In above expressions $V$ stands for volume, $T$ the temperature, $k_{\rm B}$ the Boltzmann constant, $P$ the pressure, $\kappa_T$ the isothermal compressibility, $\alpha_{P}$ the thermal expansion coefficient, $\langle N \rangle$ is the average number of particles, with averaging over the grand canonical distribution.
 
-The equation of state obtained in~\cite{KD20} reads
-\begin{equation}
-	\label{eq:eosMT}
-	Pv\beta = E_\mu(M, T) + M \bar \rho_0 + \frac{1}{2} d \bar \rho_0^2 - \frac{a_4}{24} \bar \rho_0^4.
-\end{equation}
-The quantities in the left-hand side of the equation are $P$, the pressure; $\beta = (k_{\rm B} T)^{-1}$, the inverse temperature; $k_{\rm B}$, the Boltzmann constant; $T$, the temperature; $v$, cell volume.
-The quantities in the right-hand side are, in general, functions of the temperature $T$ and the chemical potential $\mu$.
-Let us present their expressions explicitly.
+The equation of state obtained in~\cite{KD20} and rewritten in reduced quantities in~\cite{DKRP24,DKRP24arxiv} reads
+\begin{eqnarray}\label{eq:eosMT}
+	%\label{eq:eosPTM_reduced}
+	P^* & = & (1 + \tau)T^*_c \bigg[ E_\mu(M, T) 
+	\nonumber\\
+	&& + M \bar \rho_0 + \frac{1}{2} d \bar \rho_0^2 - \frac{a_4}{24} \bar \rho_0^4
+	\bigg].
+\end{eqnarray}
+Let us explicitly present the expressions entering the right-hand side of this equation.
 
@@
 	& g_4 = T_4/T_0 - g_1^4 - 6 g_1^2 g_2 - 4 g_1 g_3 - 3 g_2^2, 
-	\nonumber\\
-	& a_4 = -g_4,
 \end{align}
@@
 \end{equation}
-Here $\alpha^*=v e^{\beta_c\mu_0}$, and the parameter $p$ has the form
+Here $\alpha^*=v e^{\beta_c\mu_0}$, where $\beta_c$ is the critical value of the inverse temperature $\beta = (k_{\rm B}T)^{-1}$, and the parameter $p$ has the form
 \begin{equation}
-	p = \frac{\beta_c}{2} \Phi^{(r)}(0) [\chi_0 + A_\gamma].
+	p = \frac{1}{2T^*_c} \frac{\Phi^{(r)}(0)}{\varepsilon} [\chi_0 + A_\gamma].
 \end{equation} 
-
-The quantity $\beta_c$ denotes the critical value of the inverse temperature. In~\cite{KD20} it was found that
-\begin{equation*}
-	\varepsilon\beta_c = 0.200, \quad \frac{k_{\rm B} T_c}{\varepsilon} = 4.995.
-\end{equation*}
-We also use the reduced temperature defined as $T^* = k_{\rm B}T/\varepsilon$, and thus its critical value $T^*_c = 4.995.$
+The critical temperature found in~\cite{KD20} is $T_c^* = 4.995.$
 
@@
 \begin{equation}
-	k_{\rm B}T_c = \left(g_2 - \frac{1}{2} \frac{g_3^2}{g_4} \right) (B - 1 + \chi_0) \Phi^{(r)}(0).
+	T^*_c = \left(g_2 - \frac{1}{2} \frac{g_3^2}{g_4} \right) (B - 1 + \chi_0) \frac{\Phi^{(r)}(0)}{\varepsilon}.
 \end{equation}  
-The function $E_\mu(M, T)$ from the equation \eqref{eq:eosMT} is provided by
+The function $E_\mu(M, T)$ from the equation \eqref{eq:eosMT} is given by
 \begin{eqnarray}\label{eq:E_mu}
@@
 \begin{equation}\label{eq:ro_M}
-	M + d \bar\rho_0 - \frac{a_4}{6} \bar\rho_0^3 = 0.
+	M + d \bar\rho_0 + \frac{g_4}{6} \bar\rho_0^3 = 0.
 \end{equation}
@@
 Thus, $\bar{\rho}_0$ is a function of the temperature and the chemical potential.
-
-Let us introduce the reduced pressure
-\begin{equation}
-	\label{def:reduced_pres}
-	P^* \equiv \frac{P v}{\varepsilon}.
-\end{equation}
-Considering the equation of state~\eqref{eq:eosMT}, $P^*$ is explicitly written as
-\begin{eqnarray}
-	\label{eq:eosPTM_reduced}
-	P^* & = & (1 + \tau)T^*_c \bigg[ E_\mu(M, T) 
-	\nonumber\\
-	&& + M \bar \rho_0 + \frac{1}{2} d \bar \rho_0^2 - \frac{a_4}{24} \bar \rho_0^4
-	\bigg].
-\end{eqnarray}
-
-*****************************************************************************************
-
-End of literal copy
-
-*****************************************************************************************
 
@@
 Summing up, below the critical temperature, $M=0$ is the condition for phase coexistence.
-Substituring this condition into the EoS~\eqref{eq:eosPTM_reduced}, we obtain a phase diagram in pressure-temperature plane (see Figure~\ref{fig2}).
+Substituting this condition into the EoS~\eqref{eq:eosMT}, we obtain a phase diagram in pressure-temperature plane (see Figure~\ref{fig2}).
 We will use the same condition to obtain a continuation of the coexistence line into the supercritical region.
@@
 	\includegraphics[width=0.45\textwidth]{f2.pdf} 
-	\vskip-3mm\caption{Phase diagram in the $P^*-\tau$ plane for the cell fluid model with the interaction potential \eqref{def:mod_morse}.
-The result is based on the equation of state~\eqref{eq:eosPTM_reduced}.
+	\vskip-3mm\caption{Phase diagram in the $P^*-\tau$ plane for the cell fluid model with the interaction potential \eqref{def:mod_morse}. The result is based on the equation of state~\eqref{eq:eosMT}. 
 	}\label{fig2}
@@
 \end{equation}
-In \cite{DKRP24arxiv,DKRP24} we derived expression for the reduced isothermal compressibility $\kappa^*_T \equiv \dfrac{\varepsilon \kappa_T}{v}$  in terms of thermodynamic derivatives suitable for the reduced form of the equation of state in pressure-temperture-chemical potential terms \eqref{eq:eosPTM_reduced}
+In \cite{DKRP24arxiv,DKRP24} we derived expression for the reduced isothermal compressibility $\kappa^*_T \equiv \dfrac{\varepsilon \kappa_T}{v}$  in terms of thermodynamic derivatives suitable for the reduced form of the equation of state in pressure-temperture-chemical potential terms \eqref{eq:eosMT}
 It is expressed in terms of the reduced pressure $P^*$ and the reduced particle number density $\rho^*$ as follows
@@
 \begin{equation}
-	\rho^* \equiv \frac{\langle N \rangle}{V} v = \left(\frac{\partial (Pv\beta)}{\partial (\beta \mu)}\right)_{T,V}.
+	\rho^* \equiv \frac{\langle N \rangle}{V} v = \varepsilon\left(\frac{\partial P^*}{\partial \mu}\right)_{T,V}.
 \end{equation}
@@
 Its numerical value for parameters~\eqref{params} is $\rho^*_c = 0.978$.
-Finally, derive the isothermal compressibility in explicit form  
+Finally, the isothermal compressibility takes on the explicit form  
 \begin{eqnarray}
@@
 \end{figure}
-To identify the Widom line, the locus of the isothermal compressibility maxima is found by solving the equation
+To identify the Widom line based on the locus of the isothermal compressibility maxima, we solve the equation
 \begin{equation}\label{eq:maximize}
@@
 with respect to $M$.
-Substituting the solution of \eqref{eq:maximize} into the equation of state \eqref{def:reduced_pres}, one locates the Widom line in the pressure-temperature plane (see Figure~\ref{fig4}, dashed line), expanding the phase diagram displayed in Figure~\ref{fig2} beyond the critical point.
+Substituting the solution of \eqref{eq:maximize} into the equation of state \eqref{eq:eosMT}, one locates the Widom line in the pressure-temperature plane (see Figure~\ref{fig4}, dashed line), expanding the phase diagram displayed in Figure~\ref{fig2} beyond the critical point.
 
+Similarly to the isothermal compressibility, similar arguments apply to the thermal expansion coefficient defined as
+\begin{equation}
+	\alpha_{P} = \frac{1}{V} \left(\frac{\partial V}{\partial T}\right)_{P,N}.
+\end{equation}
+Its reduced counterpart is calculated based on EoS~\eqref{eq:eosMT} by formula
+\begin{eqnarray}
+	\alpha^*_P & = & \frac{1}{T^*_c \rho^*}
+	\left[ 
+	-\left(\frac{\partial \rho^*}{\partial \tau}\right)_{\mu}
+	\right. \nonumber\\
+	&& + \left.\left(\frac{\partial \rho^*}{\partial \mu}\right)_{T}
+	\left(\frac{\partial P^*}{\partial \tau}\right)_{\mu}
+	\left(\frac{\partial P^*}{\partial \mu}\right)^{-1}_{T} 
+	\right].
+\end{eqnarray}
+Figure~\ref{fig3b} illustrates the dependence of $\alpha^*_{P}$ on $M$ at some given values of $\tau$, at $T>T_c$.
+The thermal expansion coefficient also develops a ridge of maxima in the supercritical phase, that vanishes at $\tau > 0.35$.
+To identify the Widom line based on the locus of the thermal expansion coefficient maxima, we solve the equation
+\begin{equation}\label{eq:maximize_alpha}
+	\frac{\partial \alpha_P^*}{\partial M} = 0
+\end{equation}
+with respect to $M$.
+Substituting the solution of \eqref{eq:maximize_alpha} into the equation of state \eqref{eq:eosMT}, one locates another Widom line in the pressure-temperature plane (see Figure~\ref{fig4}, dash-dotted line).
+This is the second option for the separation line between liquid-like and gas-like fluid.
+
+The third choice for the boundary between liquid-like and gas-like regions beyond CP is based on the condition $M=0$.
 On the one hand, within the present approach, the condition $M=0$ holds along phase coexistence up to the critical point.
@@
 On the other hand, taking into account the functional form of the grand partition function~\cite{KD16,KR09}, the condition of $M=0$ is equivalent to the absence of an external field in the Ising model.
-This analogy is interesting from the perspective of mapping between the Ising model and fluids~\cite{Kulinskii10jcp,BK11}.
+This analogy is interesting from the perspective of direct mapping between the Ising model and fluids~\cite{Kulinskii10jcp,BK11}.
+In Fig.~\ref{fig4} the separation line based on this condition $M=0$ is given by the dotted curve.
 
@@
 	\vskip-3mm\caption{Complete pressure-temperature phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse}.
-Comparison of the Widom line (maxima locus of isothermal compressibility and thermal pressure) and the crossover between gas-like and liquid-like structures of the supercritical cell fluid based on combination of the equation of state~\eqref{eq:eosPTM_reduced} and the condition $M = 0$.
+Comparison of the Widom line (maxima locus of isothermal compressibility and thermal pressure) and the crossover between gas-like and liquid-like structures of the supercritical cell fluid based on combination of the equation of state~\eqref{eq:eosMT} and the condition $M = 0$.
 An inset in the right bottom corner focuses on a range of $\tau$ around its critical value $-0.01 < \tau < 0.05$.	}\label{fig4}
@@
 \section{Conclusions}
-The locus of the isothermal compressibility maxima is calculated for the cell fluid model with the modified Morse potential.
-This is considered as a good candidate for a separation line between gas-like and liquid-like regions in the supercritical phase.
-We also suggest another such line derived from the condition for the effective chemical potential $M=0$.
-This is a natural extension of the coexistence line below CP, since the same condition is used in both cases.
-A noticeable difference between the two separation lines is that the compressibility maxima end at some temperature (or pressure) value above CP, see Fig.~\ref{fig4}, while the second line extends infinitely far from CP.
-Investigation of how physical quantities change their behavior when crossing these lines will be the subject of next studies.
+The supercritical region of the phase diagram is studied for the cell fluid model with the modified Morse potential.
+Three lines are found that can be considered as good candidates for a separation boundary between gas-like and liquid-like regions in the supercritical phase.
+The first line is the locus of the isothermal compressibility maxima.
+The second line is the locus of the thermal expansion coefficient maxima.
+The third line is the location of states where the condition $M=0$ holds true.
+This third option seems to be a natural extension of the coexistence line beyond the critical point, since the same condition is used both below and above $T_c$.
+A noticeable difference between two loci of thermodynamic response functions maxima and the line obtained from $M=0$ is that the latter one is extended over the phase diagram without any limits, while the first two have their termination points.
+Investigation of how physical quantities change their behavior when crossing these lines will be the subject of our future studies.
 

### 2024-10-29 main_widom_morse.tex
change text to mentin three lines

@@
 Since our model allows for analytical expressions for many thermodynamic quantities~\cite{KD20,DKRP24}, it is straightforward to calculate and build such continuations.
-As a logical sequence of our previous works for the cell model with the modified Morse potential, in this paper we present results for two lines that can be cosidered as continuations of the coexistence line beyond the critical point in the pressure-temperature plane.
+As a logical sequence of our previous works for the cell model with the modified Morse potential, in this paper we present results for three lines that can be cosidered as continuations of the coexistence line beyond the critical point in the pressure-temperature plane.
 The first one is the locus of the isothermal compressibility maxima.
-This line, as will be shown, ends (disappears, ceases to exist) at some temperature above the critical one.
-The second one is a line that is derived from the same condition that the first-order phase transition line below the critical point - the condition on the effective chemical potential $M=0$.
+This line, as will be shown, terminates at some temperature above the critical one.
+The second line is the locus of the thermal expansion coefficient, which also terminates at some point.
+The third one is a line that is derived from the same condition that the first-order phase transition line below the critical point - the condition on the effective chemical potential $M=0$.
 This line can be extended into the supercritical region without any limits in temperature or pressure.
@@
 In Section~\ref{sec:eos}, we recall the form of the modified Morse potential, present the explicit expression for the equation of state, and illustrate pressure isotherms at temperatures below and above the critical one.
-In Section~\ref{sec:widom}, we display the Widom line as the maxima locus of the isothermal compressibility and introduce the supercritical crossover line obtained directly from the equation of state of the cell fluid model applying the proper condition on the effective chemical potential.
+In Section~\ref{sec:widom}, we develop two alternatives of the Widom line, as the maxima locus of the isothermal compressibility and as the locus of maxima of the thermal expansion coefficient.
+Then we introduce the supercritical crossover line obtained directly from the equation of state of the cell fluid model applying a condition on the effective chemical potential that is the same as is used below CP to obtain the coexistence line.
 

### 2024-10-30 main_widom_morse.tex
fixed diacritics in bib-file, tex-file: some corrections in text, add terminating P*, add keywords in eng (commented), add abstract and keywords in ukr (commented), add bibliography in UJP style after end-document

@@
 By the symmetry of the functionals for the partition functions, the condition $M=0$ in fluids is analogous to the absence of an external field in the Ising model.
+
+% cell fluid model, Morse potential, supercritical region, Widom line
+
+% Продовжується дослідження коміркової моделі плину з модифікованим потенціалом Морзе.
+Надкритичні стани розглядаються з огляду на можливість побудови межі розділення між рідиноподібними та газоподібними поведінками.
+Розраховано три різні лінії, які можна використовувати для цієї мети: лінія максимумів ізотермічної стисливості, лінія максимумів коефіцієнта теплового розширення та лінія, де ефективний хімічний потенціал дорівнює нулю, $M=0$.
+За симетрією функціоналів статистичних сум умова $M=0$ у рідинах аналогічна відсутності зовнішнього поля в моделі Ізінга.
+{\textit{К\,л\,ю\,ч\,о\,в\,і\,	с\,л\,о\,в\,а:} коміркова модель, потенціал Морзе, надкритична область, лінія Відома.}
 
@@
 This occurs because near CP, the system experiences significant fluctuations in density over increasingly large spatial scales.
-Although, the equation of state~\eqref{eq:eosPTM_reduced} is derived in the mean-field-like approximation, we established in~\cite{DKRP24arxiv,DKRP24} that the isothermal compressibility of the cell fluid model diverges at CP and consequently displays maxima in the supercritical region (see Figures~\ref{fig3}).
+Although, the equation of state~\eqref{eq:eosMT} is derived in the mean-field-like approximation, we established in~\cite{DKRP24arxiv,DKRP24} that the isothermal compressibility of the cell fluid model diverges at CP and consequently displays maxima in the supercritical region (see Figures~\ref{fig3}).
 This suggests the possibility of deriving the Widom line and expanding the phase diagram in Figure~\ref{fig2} beyond CP.
@@
 Far beyond the critical point ($\tau > 0.5$), there is no noticeable difference in $\kappa_T$ values around the effective chemical potential $M=0$.
-At $\tau = 0.2$ the compressibility maxima become almost undistinguished, they completely vanish at $\tau > 0.28$.
+At $\tau = 0.2, P^*>2.30$ the compressibility maxima become almost undistinguished, they completely vanish at $\tau > 0.28$.
 \begin{figure}[h!] 
@@
 	\caption{The reduced thermal expansion $\alpha^*_P$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$.
-The plot focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\alpha^*_T$.
+The plot focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\alpha^*_P$.
 	}
@@
 Figure~\ref{fig3b} illustrates the dependence of $\alpha^*_{P}$ on $M$ at some given values of $\tau$, at $T>T_c$.
-The thermal expansion coefficient also develops a ridge of maxima in the supercritical phase, that vanishes at $\tau > 0.35$.
+The thermal expansion coefficient also develops a ridge of maxima in the supercritical phase, that almost vanishes at $\tau > 0.35, P^*>2.75$.
 To identify the Widom line based on the locus of the thermal expansion coefficient maxima, we solve the equation
@@
 The third choice for the boundary between liquid-like and gas-like regions beyond CP is based on the condition $M=0$.
-On the one hand, within the present approach, the condition $M=0$ holds along phase coexistence up to the critical point.
+Within the present approach, the condition $M=0$ holds along phase coexistence up to the critical point.
 Thus, the implementation of the condition $M=0$ above CP seems to be a natural continuation of the coexistence line into the supercritical region.
@@
 In Fig.~\ref{fig4} the separation line based on this condition $M=0$ is given by the dotted curve.
+In the critical region ($\tau<0.02$), beyond CP, the maxima locus of the response functions and the $M=0$ based line visually merge into a single line.
 
@@
 
-	%@incollection{KKD2018,
-		%  title={Phase Transitions in a Continuum {C}urie-{W}eiss System: A Quantitative Analysis},
-		%  author={Kozitsky, Yu. and Kozlovskii, M. and Dobush, O.},
-		%  booktitle={Modern Problems of Molecular Physics},
-		%  editor={Bulavin, L. A. Chalyi, A. V.},
-		%  pages={229--251},
-		%  year={2018},
-		%  publisher={Springer},
-		%  address={Cham, Switzerland}
-
-		\bibitem{KnezEtAl2014} {\v{Z}}.
-Knez, E.
-Marko{\v{c}}i{\v{c}}, M.
-Leitgeb, M.
-Primo{\v{z}}i{\v{c}}, M.
-Knez Hrn{\v{c}}i{\v{c}}, M.
-{\v{S}}kerget, Industrial applications of supercritical fluids: A review, Energy, 77 (2014) 235-243.
-% https://doi.org/10.1016/j.energy.2014.07.044
-		\bibitem{BrykEtAl2017} T.
-Bryk, F.
-A.
-Gorelli, I.
-Mryglod, G.
-Ruocco, M.
-Santoro, T.
-Scopigno.
-Behavior of Supercritical Fluids across the “Frenkel Line”.
-The Journal of Physical Chemistry Letters 2017 8 (20), 4995-5001 %	DOI: 10.1021/acs.jpclett.7b02176
-		\bibitem{CampiEtAl2001} Campi, X., Krivine, H., Sator, N.
-Percolation line of self-bound clusters in supercritical fluids.
-Physica A: Statistical Mechanics and its Applications, 296(1-2), 24-30 (2001).
-%https://doi.org/10.1016/S0378-4371(01)00158-3
-		\bibitem{XuEtAl2005}Xu, L., Kumar, P., Buldyrev, S.
-V., Chen, S.-H., Poole, P.
-H., Sciortino, F., Stanley, H.
-E.
-Relation between the Widom line and the dynamic crossover in systems with a liquid-liquid phase transition.
-Proceedings of the National Academy of Sciences, 102(46), 16558–16562  (2005).
-doi:10.1073/pnas.0507870102
-		\bibitem{McMillanStanley2010} McMillan, P., Stanley, H. Going supercritical. Nature Phys 6, 479–480 (2010). %https://doi.org/10.1038/nphys1711
-		\bibitem{SimeoniEtAl2010} Simeoni, G., Bryk, T., Gorelli, F.
-et al.
-The Widom line as the crossover between liquid-like and gas-like behaviour in supercritical fluids.
-Nature Phys 6, 503–507 (2010).
-%https://doi.org/10.1038/nphys1683
-		\bibitem{ArtemenkoEtAl2017} Artemenko, S., Krijgsman, P., Mazur, V.
-The Widom line for supercritical fluids.
-Journal of Molecular Liquids, 238, 122-128 (2017).
-%http://dx.doi.org/10.1016/j.molliq.2017.03.107
-		\bibitem{Proctor2020Book} J.~E. Proctor, The liquid and supercritical fluid states of matter. CRC Press (2020). %[ISBN 9780429491443]
-
-		\bibitem{KozlovskiiDobush2020}  M.P. Kozlovskii and O.A. Dobush,  Ukr. J. Phys. {\bf 65} 428 (2020).
-		\bibitem{DobushEtAl2024} O.A. Dobush, M.P. Kozlovskii, R.V. Romanik, I.V. Pylyuk, arXiv preprint 2409.09786 (2024)
-		\bibitem{KozlovskiiDobush2016}  M. Kozlovskii and O. Dobush, J. Mol. Liq., {\bf 215}, 58, (2016).
-
-%		\bibitem{refb}  I.V. Pylyuk and O.A. Dobush,  Ukr. J. Phys. {\bf 65} 1080 (2020).
-%		\bibitem{refc0} I.V. Pylyuk and M.P. Kozlovskii, O.A. Dobush, M.V. Dufanets, J. Mol. Liq., {\bf 355}, 122322, (2023).
-%		\bibitem{refc1} I.V. Pylyuk and M.P. Kozlovskii, O.A. Dobush, Ukr. J. Phys., {\bf 68}, 601, (2023).
-%		\bibitem{refc}  M.P. Kozlovskii and O.A. Dobush, Condens. Matter Phys., {\bf 20}, 23501,(2017).
-%		\bibitem{refd}  M.P. Kozlovskii and I.V. Pylyuk, O.A. Dobush, Condens. Matter Phys., {\bf 21}, 43502, (2018).
-%		\bibitem{ref8}  R.C. Lincoln and K.M. Koliwad, Phys. Rev., {\bf 157}, 463, (1967).
-%		\bibitem{ref9}  J.K. Singh, J. Adhikari, and S.K. Kwak, Fluid Phase Equilibria, {\bf 248}, 1, (2006).
-
-%		\bibitem{ref16} M.V.~Fedoryuk, \textit{Asymptotic methods in analysis} in {\it Analysis I: Integral Representations and Asymptotic Methods}, eds.
-M.A.
-Evgrafov and R.V.
-Gamkrelidze  (Encyclopaedia of Mathematical Sciences 13, Springer-Verlag Berlin Heidelberg, 1989) pp.~83--191.
-
-%		\bibitem{ref11} A. J. Schultz and D. A. Kofke, Fluid Phase Equilibria, {\bf 409}, 12, (2016).
-%		\bibitem{ref12} A. J. Masters, J. Phys.: Condens. Matter, {\bf 20}, 283102, (2008).
-%		\bibitem{ref13} M. V. Ushcats, J. Chem. Phys., {\bf 140}, 234309, (2014).
-%		\bibitem{ref14} D. Pini and G. Stell, N. B. Wilding, Mol. Phys., {\bf95}, 483, (1998).
-%		\bibitem{ref15} C.-L. Lee and G. Stell, J. Hoye, J. Mol. Liq., {\bf 112}, 13, (2004).
-%		\bibitem{ref1}  Y. Kozitsky and M. Kozlovskii, O. Dobush, In \emph{Modern Problems of Molecular Physics}, Springer, 229--251 (2018).
-%		\bibitem{ref3}  M.P. Kozlovskii and O.A. Dobush, I.V. Pylyuk, Ukr. J. Phys. {\bf 62} 865 (2017).
-%		\bibitem{ref4}  I. Yukhnovskii and V. Kolomiets, I. Idzyk, Condens. Matter Phys., {\bf 16}, 23604, (2013).
-%		\bibitem{ref5}  I.R. Yukhnovskii, Condens. Matter Phys., {\bf 17}, 43001, (2014).
-
-%		\bibitem{ref10} I.R. Yukhnovskii and M.P. Kozlovskii, I.V. Pylyuk, Phys. Rev. B, {\bf 66}, 134410, (2002).
-%		\bibitem{ref17} D. Pini and G. Stell, N. B. Wilding, Mol. Phys., \textbf{95}, 483, (1998).
-%		\bibitem{ref18} J.-M. Caillol, Mol. Phys., {\bf 104}, 1931, (2006).
-%		\bibitem{ref19} I.R Yukhnovskii., Theoret. and Math. Phys., \textbf{194}, 224, (2018).
-%		\bibitem{ref20} A. Oleinikova and L. Bulavin, V. Pipich, Chem. Phys. Let., {\bf 278}, 121, (1997).
-%		\bibitem{ref21} G. Brunner, J. Supercrit. Fluids, {\bf 47}, 373, (2009).
-%		\bibitem{ref22} G. Brunner, J. Supercrit. Fluids, {\bf 47}, 382, (2009).
-
-%	\bibitem{ref1} 
-%	\bibitem{ref2}
-%	\bibitem{ref3}
-%	\bibitem{ref4} 
-%	\bibitem{ref5}
-%	\bibitem{ref6}
-%	\bibitem{ref7}
-%	\bibitem{ref8}
-%	\bibitem{ref9}
-%	\bibitem{ref10}
-%	\bibitem{ref11}
-%	\bibitem{ref12}
-%	\bibitem{ref13}
-%	\bibitem{ref14}
-%	\bibitem{ref15}
-%	\bibitem{ref16}
-%	\bibitem{ref17}
+	%1
+	\bibitem{KMLetal14} {\v{Z}}.~Knez, E.~Marko{\v{c}}i{\v{c}}, M.~Leitgeb, M.~Primo{\v{z}}i{\v{c}}, M.~Knez Hrn{\v{c}}i{\v{c}}, M.~{\v{S}}kerget.
+	\href{https://doi.org/10.1016/j.energy.2014.07.044}{Industrial applications of supercritical fluids: A review}. \textit{Energy} \textbf{77} 235 (2014) [doi: 10.1016/j.energy.2014.07.044]
+	%2
+	\bibitem{FW69} M.E.~Fisher, B.~Widom.
+	\href{https://doi.org/10.1063/1.1671624}{Decay of Correlations in Linear Systems}. \textit{J. Chem. Phys.} \textbf{50} 3756 (1969) [doi: 10.1063/1.1671624]
+	%3
+	\bibitem{BGMetal17} T.~Bryk, F.A.~Gorelli, I.~Mryglod, G.~Ruocco, M.~Santoro, T.~Scopigno.
+	\href{https://doi.org/10.1021/acs.jpclett.7b02176}{Behavior of Supercritical Fluids across the “Frenkel Line”}. \textit{J. Chem. Phys. Lett.} \textbf{8} 4995 (2017) [doi: 10.1021/acs.jpclett.7b02176]
+	%4
+	\bibitem{CKS01} X.~Campi, H.~Krivine, N.~Sator.
+	\href{https://doi.org/10.1016/S0378-4371(01)00158-3}{Percolation line of self-bound clusters in supercritical fluids}. \textit{Physica A} \textbf{296} 24 (2001) [doi: 10.1016/S0378-4371(01)00158-3] 
+	%5
+	\bibitem{XKBetal05} L.~Xu, P.~Kumar, S.V.~Buldyrev, S.-H.~Chen, P.H.~Poole, F.~Sciortino, H.E.~Stanley.
+	\href{https://doi.org/10.1073/pnas.0507870102}{Relation between the Widom line and the dynamic crossover in systems with a liquid-liquid phase transition}.
+\textit{PNAS} \textbf{102} 16558 (2005) [doi: 10.1073/pnas.0507870102]
+	%6
+	\bibitem{MS10} P.~McMillan, H.~Stanley.
+	\href{https://doi.org/10.1038/nphys1711}{Going supercritical}. \textit{Nature Phys.} \textbf{6} 479 (2010) [doi: 10.1038/nphys1711]
+	%7
+	\bibitem{AKM17} S.~Artemenko, P.~Krijgsman, V.~Mazur. 
+	\href{https://doi.org/10.1016/j.molliq.2017.03.107}{The Widom line for supercritical fluids}. \textit{J. Mol. Liq.} \textbf{238} 122 (2017) [doi: 10.1016/j.molliq.2017.03.107]
+	%8
+	\bibitem{Proctor20} J.E.~Proctor.
+	\href{https://doi.org/10.1201/9780429491443}{The liquid and supercritical fluid states of matter}. (CRC Press, 2020)  [ISBN: 9780429491443]
+	%9
+	\bibitem{KD20} M.~Kozlovskii, O.~Dobush.
+	\href{https://ujp.bitp.kiev.ua/index.php/ujp/article/view/2019631}{Phase behavior of a cell fluid model with modified Morse potential.} \textit{Ukr.
+J.
+Phys.} \textbf{65}, 428 (2020) [doi: 10.15407/ujpe65.5.428]
+	%10
+	\bibitem{DKRP24} O.A.~Dobush, M.P.~Kozlovskii, R.V.~Romanik, I.V.~Pylyuk.
+	\href{https://doi.org/10.48550/arXiv.2409.09786}{Thermodynamic response functions in a cell fluid model}. \textit{Ukr. J. Phys} - in print [doi: 10.48550/arXiv.2409.09786]
+	%11
+	\bibitem{DKRP24arxiv} O.A.~Dobush, M.P.~Kozlovskii, R.V.~Romanik, I.V.~Pylyuk.
+	\href{https://doi.org/10.48550/arXiv.2409.09786}{Thermodynamic response functions in a cell fluid model}. \textit{arXiv preprint} 2409.09786 (2024) [doi: 10.48550/arXiv.2409.09786]
+	%12
+	\bibitem{Morse29} P.~M.~Morse. 
+	\href{https://link.aps.org/doi/10.1103/PhysRev.34.57}{Diatomic molecules according to the wave mechanics.
+II.
+Vibrational levels.} \textit{Phys.
+Rev.} \textbf{34}, 57 (1929) [doi: 10.1103/PhysRev.34.57]
+	%13
+	\bibitem{SBGetal10} G.G.~Simeoni, T.~Bryk, F.A.~Gorelli, M.~Krisch, G.~Ruocco, M.~Santoro, T.~Scopigno.
+	\href{https://doi.org/10.1038/nphys1683}{ The Widom line as the crossover between liquid-like and gas-like behaviour in supercritical fluids}.
+\textit{Nature Phys.} \textbf{6} 503 (2010) [doi: 10.1038/nphys1683]
+	%14
+	\bibitem{KozlovskiiDobush2016} M.~Kozlovskii, O.~Dobush.
+	\href{https://doi.org/j.molliq.2015.12.018}{Representation of the grand partition function of the cell model: The state equation in the mean-field approximation.} \textit{J.
+Mol.
+Liq.} \textbf{215}, 58 (2016) [doi: j.molliq.2015.12.018]
+	%15
+	\bibitem{KR09} M.P~Kozlovskii, R.V.~Romanik.
+	\href{https://doi.org/10.30970/jps.13.4007}{The equation of state of a three-dimensional {Ising}-like system}. \textit{Journal of Physical Studies} \textbf{13} 4007 (2009) [doi: 10.30970/jps.13.4007]
+	%16
+	\bibitem{Kulinskii10jcp} V.L.~Kulinskii.
+	\href{https://doi.org/10.1063/1.3457943}{Global isomorphism between the {Lennard-Jones} fluids and the {Ising} model}. \textit{J. Chem. Phys.} \textbf{133} 034121 (2010) [doi: 10.1063/1.3457943]
+	%17
+	\bibitem{BK11} L.A.~Bulavin, V.L.~Kulinskii.
+	\href{https://doi.org/10.1021/jp201872f}{Unified Picture for the Classical Laws of {B}atschinski and the Rectilinear Diameter for Molecular Fluids}.
+\textit{J.
+Phys.
+Chem.
+B} \textbf{115} 6061 (2011) [doi: 10.1021/jp201872f]
 
@@
 %\end{document}
-
-	The latter expression links the pressure $P$, the temperature $T$ ($\beta = 1/T$ is the inverse temperature), the cell's volume $v$ and the effective chemical potential $M$.
-In general $M$ is a function of the real chemical potential $\mu$:
-\begin{align}
-	&	M = \frac{\tilde\mu}{W(0)} + g_1 + n_c \tilde D(0) - \frac{1}{6} \frac{g_3^3}{g_4^2}, \\
-	&	\tilde\mu=\mu-\mu^*(1+\tau),
-\end{align}
-here $\mu^*$ is some positive constant, and $\tau$ is the relative temperature $\tau = (T - T_c) / T_c$,
-$T_c$ is the critical temperature. $W(0)$ is what we call an effective interaction potential\footnote{\textbf{explain} what is the effective interaction potential in a few words} at ${\bf k} = 0$:
-\begin{equation}
-	W(0) = \Phi^{(r)}(0) \left[ B - 1 + \chi_0 + \tau (\chi_0 + A_\gamma) \right].
-\end{equation}
-Here $\chi_0$ is a constant parameter, the quantities $B$ and $A_\gamma$ reads
-\begin{align} 
-	& B = 2 \gamma^3 e^{(1-\gamma)/\alpha},\nonumber \\
-	& A_\gamma = A e^{(n_0-\gamma)/\alpha} \left( \gamma / n_0\right)^3.
-\end{align}
-The coefficients $g_n$ are given by the formulas:
-\begin{align}
-	& g_0 = \ln T_0, \qquad g_1 = T_1/T_0, \qquad g_2 = T_2/T_0 - g_1^2,  \nonumber \\
-	& g_3 = T_3/T_0 - g_1^3 - 3g_1 g_2, \quad \qquad  g_4 = T_4/T_0 - g_1^4 - 6 g_1^2 g_2 - 4 g_1 g_3 - 3 g_2^2,
-\end{align}
-where $T_n(p,\alpha^*)$ are the following special functions
-\begin{equation}
-	T_n(p,\alpha^*) = \sum_{m=0}^{\infty} \frac{(\alpha^*)^m}{m!} m^n e^{pm^2}.
-\end{equation}
-Here $\alpha^*=v e^{\beta_c\mu^*}$, and the parameter $p$ has the form
-\begin{equation}
-	p = \frac{\beta_c}{2} \Phi^{(r)}(0) [\chi_0 + A_\gamma].
-\end{equation} 
-Note that we denote $\beta_c$ as the inverse critical temperature.
-Since $p$ fails to be the function of temperature, the coefficients $g_n$ are also temperature-free.
-The quantity $\tilde D(0)$ which has already appeared in both equations \eqref{eq:eos} and \eqref{3d2} is a function of temperature
-\begin{equation}
-	\tilde D(0) = g_2 - \frac{1}{2} \frac{g_3^2}{g_4} - \frac{1}{\beta W(0)},
-\end{equation}
-which equals zero at $T=T_c$. This circumstance gives us an expression for determining the critical temperature 
-\begin{equation}
-	T_c = \left(g_2 - \frac{1}{2} \frac{g_3^2}{g_4} \right) (B - 1 + \chi_0) \Phi^{(r)}(0).
-\end{equation}    
-The function $E_\mu(M, T)$ from the equation \eqref{eq:eos} is provided by
-\begin{align}
-	& E_\mu (M, T) = - \frac{1}{2 N_v} \ln (2\pi \beta W(0)) +  g_0 - \frac{\beta W(0)}{2} \!\!
-\left( \!
-\frac{\tilde\mu}{W(0)} \!
-\right)^{\!
-2} \!\!\!
-- \frac{g_3}{g_4} {M} \!
-- \frac{g_3^2}{2 g_4^2}  \tilde D(0) - \frac{1}{24} \frac{g_3^4}{g_4^3}.
-\end{align}
-Note that for all $ \tau> 0 $ there are no restrictions on the value of $ \bar \rho_0 $, which is a function of the temperature and the chemical potential
-\begin{equation}\label{2d18}
-	\bar \rho_0 = \left(- \frac{3\bar{M}}{g_4} + \sqrt{Q_t}\right)^{1/3} - \left(  \frac{3\bar{M}}{g_4} + \sqrt{Q_t} \right)^{1/3} \!\!\!\! ,
-\end{equation}
-where
-\begin{equation}
-	Q_t = \left(  \frac{2\tilde D(0)}{g_4}\right)^3 + \left( -\frac{3\bar{M}}{g_4}\right)^2, \qquad g_4<0.
-\end{equation}  
-...
-\begin{equation}\label{3d3}
-	\bar n = n_g - \bar{M} + \tilde g_2 \gamma_\tau \bar \rho_0,
-\end{equation}
-where
-\begin{equation}\label{3d4}
-	n_g = g_1 + n_c \tilde g_2 - \frac{1}{6} g_3^3 / g_4^2.
-\end{equation}
-...

### 2024-10-30 kozlovskii_dobush_romanik_ujp_UTF8.tex
add UJP styles

\documentclass[fleqn,twoside,twocolumn,nofootinbib,showkeys]{revtex4} % Specifies the document class %,unsortedaddress
%\usepackage[sec]{ujp} % \usepackage[cyr]{ujp} for cyrillic
\usepackage[sec,doi]{ujp_UTF8} % \usepackage[cyr]{ujp_UTF8} for cyrillic

\usepackage{orcidlink} % to insert a hyperlinked ORCiD logo
\usepackage{amssymb}
\usepackage{physics}
%\usepackage{cite} % may be useful for different formats of numerical citation
\usepackage{hyperref} % for hyperlinks in references
\usepackage[nameinlink]{cleveref}
%\usepackage{caption} % for \caption*
\usepackage{amsmath}    % for "\eqref" macro
\usepackage{graphicx}

\begin{document}
	\title[Supercritical crossover lines in a cell fluid model]%колонтитул
	{Supercritical crossover lines in a cell fluid model}%
	\author{M.P.~Kozlovskii}%1
	\author{O.A.~Dobush}
	\author{R.V.~Romanik}
	%\author{I.V.~Pylyuk}
	\affiliation{Institute for Condensed Matter Physics,\\ Nat. Acad. of Sci. of Ukraine}%інститут 1
	\address{1, Svientsitskii Str., Lviv 79011, Ukraine}%адреса 1
	\email{mpk@icmp.lviv.ua}%e-mail 1
	\email{dobush@icmp.lviv.ua}
	\email{romanik@icmp.lviv.ua}
	%\email{piv@icmp.lviv.ua}

	\udk{№ УДК/UDC} \razd{\secv}

	\autorcol{O.A.\hspace*{0.7mm}Dobush, M.P.\hspace*{0.7mm}Kozlovskii, R.V.\hspace*{0.7mm}Romanik, I.V.\hspace*{0.7mm}Pylyuk}%

	\setcounter{page}{1}%

	\begin{abstract}
		A cell fluid model with a modified Morse potential is studied. The supercritical states are considered with respect to a possibility to build a separation boundary between liquid-like and gas-like bahaviors. Three different lines are calculated that can be used for this purpose: the locus of the isothermal compressibility maxima, the locus of the thermal expansion coefficient maxima, and the line where the effective chemical potential is zero, $M=0$. By the symmetry of the functionals for the partition functions, the condition $M=0$ in fluids is analogous to the absence of an external field in the Ising model.
	\end{abstract}

	\keywords{Cell model, Morse potential, Thermodynamic response functions}

	\maketitle

	\input{./paper_text.tex}
	\bibliographystyle{gost2008}
	\bibliography{extracted}
	%\bibliography{extracted}

	\vspace*{-5mm}\rezume{%Резюме укр. мовою
		М.П.~Козловський, О.А.~Добуш, Р.В.~Романік}{Надкритичні лінії переходу в комірковій моделі плину} {В роботі проведено... .}{\textit{К\,л\,ю\,ч\,о\,в\,і\,
			с\,л\,о\,в\,а:} коміркова модель, потенціал Морзе, термодинамічні функції відгуку.}

\end{document}

### 2024-10-30 main_widom_morse.tex
add UJP styles

@@
 
-\section{Introduction}
-A supercritical fluid is a state of matter that occurs when a substance is subjected to temperatures and pressures above its critical point (CP).
-Below CP the substance exhibits a liquid-gas coexistence.
-At CP, the distinction between the liquid and gas phases of a substance disappears, and beyond this point, the substance exists in a single continuous supercritical state.
-Despite the absence of first-order phase transitions above the critical point, fluids exhibit distinct liquid-like and gas-like behavior under supercritical conditions, and display a unique mix of properties.
-With densities comparable to liquids they possess low viscosity and high diffusivity like gases resulting in high solvent power.
-Moreover, the supercritical properties are instantly tunable by changing pressure and temperature.
-This is the reason why supercritical fluids have diverse applications across various industries \cite{KMLetal14}.
-
-Although the traditional liquid-gas phase distinction does not apply in the supercritical region, scientists continue to search for crossover regions or definitive lines that distinguish liquid-like and gas-like behaviors in the phase diagram.
-Naturally, such boundaries originate from CP and may either vanish away from it or extend infinitely far.
-Among known approaches to this problem are the Widom line, which represents the locus of correlation length maxima~\cite{FW69}; the Frenkel line~\cite{BGMetal17}, focusing on the dynamics of molecular motion; and the percolation line~\cite{CKS01}, related to spatial structures and marking the threshold where clusters of fluid or density fluctuations form a continuous, system-spanning network.
-
-In the present work, we focus on the Widom line \cite{XKBetal05,MS10,AKM17}, a concept based on thermodynamic response functions that represents a conditional boundary between liquid-like and gas-like regions in supercritical fluids.
-Thus, the Widom line \--- the locus of correlation length maxima emanating from CP~\cite{XKBetal05} \--- is treated as the continuation of the coexistence line above CP.
-Moreover, some thermodynamic response functions, such as isobaric heat capacity and isothermal compressibility, also exhibit maxima in the supercritical region.
-In the critical region beyond the critical point, the maxima lines of the correlation length and these response functions virtually merge into a single line~\cite{AKM17} and then diverge from each other as the temperature increases.
-Therefore, in the critical region, the Widom line is viewed as the locus of maxima of thermodynamic response functions~\cite{Proctor20}.
-
-In our previous work~\cite{KD20}, the equation of state for a many-particle system interacting via a modified Morse potential was developed in the framework of a cell model.
-Recently~\cite{DKRP24,DKRP24arxiv}, the thermodynamic response functions, namely the isothermal compressibility, the thermal pressure coefficient, and the thermal expansion coefficient, have been calculated for the same model system.
-The isothermal compressibility and thermal expansion coefficient both develop a maximum in some region beyond critical point.
-On the one hand, such response function maxima, as mentioned above, are considered as good candidates for crossover lines between gas-like and liquid-like behaviors in the supercritical region.
-On the other hand, they are also considered as natural continuation of the coexistence line.
-Since our model allows for analytical expressions for many thermodynamic quantities~\cite{KD20,DKRP24}, it is straightforward to calculate and build such continuations.
-As a logical sequence of our previous works for the cell model with the modified Morse potential, in this paper we present results for three lines that can be cosidered as continuations of the coexistence line beyond the critical point in the pressure-temperature plane.
-The first one is the locus of the isothermal compressibility maxima.
-This line, as will be shown, terminates at some temperature above the critical one.
-The second line is the locus of the thermal expansion coefficient, which also terminates at some point.
-The third one is a line that is derived from the same condition that the first-order phase transition line below the critical point - the condition on the effective chemical potential $M=0$.
-This line can be extended into the supercritical region without any limits in temperature or pressure.
-
-In Section~\ref{sec:eos}, we recall the form of the modified Morse potential, present the explicit expression for the equation of state, and illustrate pressure isotherms at temperatures below and above the critical one.
-In Section~\ref{sec:widom}, we develop two alternatives of the Widom line, as the maxima locus of the isothermal compressibility and as the locus of maxima of the thermal expansion coefficient.
-Then we introduce the supercritical crossover line obtained directly from the equation of state of the cell fluid model applying a condition on the effective chemical potential that is the same as is used below CP to obtain the coexistence line.
-
-\section{\label{sec:eos} The equation of state}
-
-In the present work, we utilize the results obtained in~\cite{KD20,DKRP24arxiv,DKRP24}.
-In~\cite{KD20}, the phase space of collective variables was used to calculate the grand partition function of a cell fluid model in the zero-mode approximation, leading to explicit forms of the equation of state (EoS) in both pressure-temperature-chemical potential and pressure-temperature-density terms.
-These EoSs are applicable across a wide range of temperatures below and above the CP.
-Furthermore, in~\cite{DKRP24arxiv,DKRP24}, we have used these equations as the basis for calculating and graphically representing the thermodynamic response functions \--- specifically, the isothermal compressibility, thermal pressure coefficient, and thermal expansion \--- of the cell fluid model with the modified Morse potential in the supercritical region.
-
-The modified Morse potential used to describe the interaction between particles is given by 
-\begin{eqnarray}
-	\label{def:mod_morse}
-	U(r) & = & \varepsilon C_{H} \left[A {\rm e}^{-n_0(r-R_0)/\alpha} \right.
-	\nonumber\\
-	&& + \left. {\rm e}^{-\gamma(r - R_0)/\alpha} -2{\rm e}^{-(r-R_0)/\alpha}\right],
-\end{eqnarray}
-where $R_0$ is the coordinate of the potential minimum, $\alpha$ is an effective range of interaction, $\gamma$ and $n_0$ are parameters of the model.
-Other two constants $C_{H}$ and $A$ are expressed via $\gamma$ and $n_0$ as follows
-\begin{equation}
-	C_{H} = \frac{n_0}{n_0 + \gamma - 2}, \quad A = \frac{2 - \gamma}{n_0},
-\end{equation}
-where $\varepsilon$ is the depth of the potential well at $r=R_0$.
-Applying $\gamma=2$ reduces $U(r)$ to the ordinary Morse potential~\cite{Morse29}.
-For a more detailed discussion of such modified Morse potential, see Sections~1 and~2 in~\cite{KD20}, and Section~2 in~\cite{DKRP24arxiv}.
-
-We use $\varepsilon$ and $v$ as natural units for energy and volume.
-Therefore, the following reduced quantities are defined: the reduced density $\rho^*=\rho v$, where $\rho = \langle N \rangle /V$; reduced temperature $T^* = k_{\rm B}T/\varepsilon$; reduced pressure $P^* = Pv/\varepsilon$; reduced isothermal compressibility $\kappa^*_{T} = \varepsilon \kappa_T/v$; and the reduced thermal expansion coefficient $\alpha^*_{P} = \varepsilon \alpha_{P} / k_{\rm B}$.
-In above expressions $V$ stands for volume, $T$ the temperature, $k_{\rm B}$ the Boltzmann constant, $P$ the pressure, $\kappa_T$ the isothermal compressibility, $\alpha_{P}$ the thermal expansion coefficient, $\langle N \rangle$ is the average number of particles, with averaging over the grand canonical distribution.
-
-The equation of state obtained in~\cite{KD20} and rewritten in reduced quantities in~\cite{DKRP24,DKRP24arxiv} reads
-\begin{eqnarray}\label{eq:eosMT}
-	%\label{eq:eosPTM_reduced}
-	P^* & = & (1 + \tau)T^*_c \bigg[ E_\mu(M, T) 
-	\nonumber\\
-	&& + M \bar \rho_0 + \frac{1}{2} d \bar \rho_0^2 - \frac{a_4}{24} \bar \rho_0^4
-	\bigg].
-\end{eqnarray}
-Let us explicitly present the expressions entering the right-hand side of this equation.
-
-First, the quantity $M$ depends linearly on the chemical potential
-\begin{align}\label{chem_pot}
-	&	M = \frac{\tilde\mu}{W(0)} + g_1 - \frac{g_3}{g_4} d - \frac{1}{6} \frac{g_3^3}{g_4^2}, \\
-	&	\tilde\mu=\mu-\mu_0(1+\tau),
-\end{align}
-where $\mu_0$ is some positive constant, $\tau$ is the relative temperature $\tau = (T - T_c) / T_c$, $T_c$ is the critical temperature. We will call $M$ the effective chemical potential.
-
-The quantity $W(0)$ is expressed via parameters of the potential~(\ref{def:mod_morse}) as follows
-\begin{equation}
-	W(0) = \Phi^{(r)}(0) \left[ B - 1 + \chi_0 + \tau (\chi_0 + A_\gamma) \right],
-\end{equation}
-where
-\begin{align*} 
-	& B = 2 \gamma^3 e^{(1-\gamma)R_0/\alpha},
-	\nonumber \\
-	& A_\gamma = A e^{(n_0-\gamma)R_0/\alpha} \left( \gamma / n_0\right)^3, 
-\end{align*}
-and $\Phi^{(r)}(0)$ is the Fourier transform of the repulsive part of the potential at $\abs{\vb k}=0$
-\begin{equation*}
-	\Phi^{(r)}(0) = \varepsilon C_H 8\pi {\rm e}^{\gamma R_0/\alpha} \left(\frac{\alpha}{\gamma R_0}\right)^3.
-\end{equation*}
-
-The parameter $\chi_0$ is used in~\cite{KD20} to single out a contribution in the Fourier transform of the potential that is treated as a reference system defined in the reciprocal space, and is selected as $\chi_0 = 0.07$~\cite{KD20}.
-
-The coefficients $g_n$ are given by the formulas:
-\begin{align}
-	& g_0 = \ln T_0, \qquad g_1 = T_1/T_0, \qquad g_2 = T_2/T_0 - g_1^2,  
-	\nonumber \\
-	& g_3 = T_3/T_0 - g_1^3 - 3g_1 g_2, 
-	\nonumber\\
-	& g_4 = T_4/T_0 - g_1^4 - 6 g_1^2 g_2 - 4 g_1 g_3 - 3 g_2^2, 
-\end{align}
-where $T_n(p,\alpha^*)$ are the following special functions
-\begin{equation}
-	T_n(p,\alpha^*) = \sum_{m=0}^{\infty} \frac{(\alpha^*)^m}{m!} m^n {\rm e}^{-pm^2}.
-\end{equation}
-Here $\alpha^*=v e^{\beta_c\mu_0}$, where $\beta_c$ is the critical value of the inverse temperature $\beta = (k_{\rm B}T)^{-1}$, and the parameter $p$ has the form
-\begin{equation}
-	p = \frac{1}{2T^*_c} \frac{\Phi^{(r)}(0)}{\varepsilon} [\chi_0 + A_\gamma].
-\end{equation} 
-The critical temperature found in~\cite{KD20} is $T_c^* = 4.995.$
-
-Since $p$ is independent of temperature, the coefficients $g_n$ are also independent of temperature.
-The numerical values for other coefficients used in this paper are the same as those in~\cite[eqs.~(5), (23), and~(24)]{KD20}:
-\begin{eqnarray}
-	\label{params}
-	\chi_0 = 0.07, & \quad \gamma = 1.65, \nonumber\\
-	n_0 = 1.521, & \quad R_0/\alpha = 2.9544, \nonumber\\
-	\alpha^* = 5.0 & \quad p = 1.0.
-\end{eqnarray}
-
-The quantity $d$ entering equations~\eqref{eq:eosMT} and~\eqref{chem_pot} is a function of temperature
-\begin{equation}
-	\label{def:D0}
-	d = g_2 - \frac{1}{2} \frac{g_3^2}{g_4} - \frac{1}{\beta W(0)}.
-\end{equation}
-The condition $d = 0$ defines the critical temperature~\cite{KD20}
-\begin{equation}
-	T^*_c = \left(g_2 - \frac{1}{2} \frac{g_3^2}{g_4} \right) (B - 1 + \chi_0) \frac{\Phi^{(r)}(0)}{\varepsilon}.
-\end{equation}  
-The function $E_\mu(M, T)$ from the equation \eqref{eq:eosMT} is given by
-\begin{eqnarray}\label{eq:E_mu}
-	E_\mu (M, T) & = & - \frac{\ln (2\pi \beta W(0))}{2 N_v}  
-	\nonumber\\
-	&& +  g_0 - \frac{\beta W(0)}{2} 
-	\left(\frac{\tilde\mu}{W(0)} \right)^{2} 
-	\nonumber\\
-	&& - \frac{g_3}{g_4} {M} \! - \frac{g_3^2}{2 g_4^2}  d - \frac{1}{24} \frac{g_3^4}{g_4^3}. 
-\end{eqnarray}
-Here the quantity $N_v$ defines the number of cubic cells in volume $V$ for the initial model.
-In the thermodynamic limit, $N_v \to \infty$, and thus, the first term can be neglected.
-The term $\tilde{\mu}/W(0)$ can be expressed in terms of $M$ using~\eqref{chem_pot}.
-The temperature and the inverse temperature can always be expressed in terms of the reduced temperature and a corresponding critical value:
-\begin{equation*}
-	T = T_c(1+\tau), \quad \beta = \beta_c (1 + \tau)^{-1}.
-\end{equation*}
-
-The quantity $\bar{\rho}_0$ is a solution to the following cubic equation
-\begin{equation}\label{eq:ro_M}
-	M + d \bar\rho_0 + \frac{g_4}{6} \bar\rho_0^3 = 0.
-\end{equation}
-For any $\tau > 0$, the latter equation has one real root
-\begin{equation}\label{eq:ro_MT}
-	\bar \rho_0 = \left(- \frac{3 M}{g_4} + \sqrt{Q_t}\right)^{1/3} - \left(  \frac{3 M}{g_4} + \sqrt{Q_t} \right)^{1/3},
-\end{equation}
-where
-\begin{equation}
-	Q_t = \left(  \frac{2d}{g_4}\right)^3 + \left( -\frac{3 M}{g_4}\right)^2, \qquad g_4<0.
-\end{equation}
-Thus, $\bar{\rho}_0$ is a function of the temperature and the chemical potential.
-
-\begin{figure}[t!]
-	\centering
-	\includegraphics[width=0.45\textwidth]{f1.pdf} 
-	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the effective chemical potential $M$. Black lines correspond to $T \geq T_c$. Grey lines correspond to $T < T_c$. 
-	}\label{fig1}
-\end{figure}
-Figure~\ref{fig1} illustrates the relationship between the reduced pressure $P^*$ and the effective chemical potential $M$ for various relative temperature values $\tau$.
-Below the critical temperature ($\tau < 0$), each іsotherm contains a kink (three bottom lines in Figure~\ref{fig1}).
-This form indicates a first-order phase transition in the system.
-Within our approach~\cite{KD20} at $T<T_c$, negative values of the effective chemical potential $M$ imply the system is in a gaseous state.
-The condition $M=0$ is a point of phase coexistence.
-Positive values of the effective chemical potential correspond to the liquid state.
-In both cases, the pressure tends to increase with $M$, but the slope of the liquid isotherm is much steeper.
-At the critical temperature, the pressure is a smooth curve (black dotted line in Figure~\ref{fig1}) with an inflection point at $M=0$.
-Thus, at the critical point coordinates are defined by $M=0, \tau=0$, $P^*_c = 1.606$.
-Beyond the critical temperature, the pressure behaves as a monotonically increasing function, and no phase transitions are observed.
-Summing up, below the critical temperature, $M=0$ is the condition for phase coexistence.
-Substituting this condition into the EoS~\eqref{eq:eosMT}, we obtain a phase diagram in pressure-temperature plane (see Figure~\ref{fig2}).
-We will use the same condition to obtain a continuation of the coexistence line into the supercritical region.
-\begin{figure}[t!]
-	\centering
-	\includegraphics[width=0.45\textwidth]{f2.pdf} 
-	\vskip-3mm\caption{Phase diagram in the $P^*-\tau$ plane for the cell fluid model with the interaction potential \eqref{def:mod_morse}. The result is based on the equation of state~\eqref{eq:eosMT}. 
-	}\label{fig2}
-\end{figure}
-
-\section{\label{sec:widom} Crossover lines between gas-like and liquid-like behavior of the supercritical fluid.}
-
- In the supercritical region, substances no longer behave distinctly as either liquids or gases but exist as a single, continuous phase without a sharp liquid-gas boundary.
-However, fluids can still exhibit liquid-like and gas-like behaviors beyond the critical temperature.
-A way to represent a conditional boundary between these regions in supercritical fluids is the Widom line \--- a continuation of the coexistence line based on the locus of maxima of certain thermodynamic response functions~\cite{XKBetal05,MS10,SBGetal10}.
-This occurs because near CP, the system experiences significant fluctuations in density over increasingly large spatial scales.
-Although, the equation of state~\eqref{eq:eosMT} is derived in the mean-field-like approximation, we established in~\cite{DKRP24arxiv,DKRP24} that the isothermal compressibility of the cell fluid model diverges at CP and consequently displays maxima in the supercritical region (see Figures~\ref{fig3}).
-This suggests the possibility of deriving the Widom line and expanding the phase diagram in Figure~\ref{fig2} beyond CP.
- In what follows we consider this in more detail.
-
-The isothermal compressibility $\kappa_T$ is defined by
-\begin{equation}
-	\label{def:isotherm_compres}
-	\kappa_T = -\frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_{T, N}.
-\end{equation}
-In \cite{DKRP24arxiv,DKRP24} we derived expression for the reduced isothermal compressibility $\kappa^*_T \equiv \dfrac{\varepsilon \kappa_T}{v}$  in terms of thermodynamic derivatives suitable for the reduced form of the equation of state in pressure-temperture-chemical potential terms \eqref{eq:eosMT}
-It is expressed in terms of the reduced pressure $P^*$ and the reduced particle number density $\rho^*$ as follows
-\begin{eqnarray}
-	\label{eq:kappa_star_m1}
-	\kappa^*_T & = & \frac{\varepsilon}{{\rho^*}^2} \left(\frac{\partial \rho^*}{\partial \mu}\right)_T
-	\\
-	\label{eq:kappa_star_m}
-	& = & \frac{\varepsilon^2}{{\rho^*}^2} \left(\frac{\partial^2 P^*}{\partial \mu^2}\right)_T.
-\end{eqnarray}
-The reduced particle number density $\rho^*$ is found by 
-\begin{equation}
-	\rho^* \equiv \frac{\langle N \rangle}{V} v = \varepsilon\left(\frac{\partial P^*}{\partial \mu}\right)_{T,V}.
-\end{equation}
-Taking explicit derivatives, we arrive at
-\begin{equation}\label{eq:density}
-	\rho^* = \rho^*_c - M + \frac{ \bar \rho_0}{\beta W(0)}.
-\end{equation}
-The quantity $\rho^*_c$ in the equation \eqref{eq:density} is the critical density~\cite{KD20,DKRP24arxiv,DKRP24}
-\begin{eqnarray}\label{eq:crit_dens}
-	\rho^*_c & = &  g_1 - \frac{g_2 g_3}{g_4} + \frac{g_3^3}{3g_4^2}.
-\end{eqnarray}
-Its numerical value for parameters~\eqref{params} is $\rho^*_c = 0.978$.
-Finally, the isothermal compressibility takes on the explicit form  
-\begin{eqnarray}
-	\kappa_T^* & = & \frac{\epsilon}{\rho^{*2} W(0)} \left\{   \frac{Q_t^{-1/2}}{\beta W(0) g_4} \left[ \bar \rho_0 - 2 \left( - \frac{3M}{g_4} + \sqrt{Q_t} \right)^{\frac{1}{3}} \right] -1 \right\}.
-\end{eqnarray}
-\begin{figure}[h!] 
-%	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
-	\centering \includegraphics[width=0.46\textwidth]{f3bb.pdf}
-	%\includegraphics[width=\column]{f2b.pdf}
-	%\includegraphics[width=\column]{f2c.pdf}
-	\vskip-3mm
-	\caption{The reduced isothermal compressibility $\kappa^*_T$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$.
-The plot focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\kappa^*_T$.
-	}
-	\label{fig3}
-\end{figure}
-Behavior of the reduced isothermal compressibility versus effective chemical potential at various fixed values of the reduced temperature is represented graphically in Figure~\ref{fig3} (narrow range around $M=0$).
-Right at the critical point the compressibility diverges.
-As a result, there is still a "memory" of the critical point, where $\kappa_T$ reaches local maxima.
-However, once the system moves further into the supercritical region, these maxima become less pronounced, and the compressibility decreases.
-Below the critical temperature, negative values of $M$ correspond to gaseous densities, while positive values of $M$ represent liquid densities.
-Above the critical point, the fluid becomes continuous.
-However, Figure~\ref{fig3} clearly shows significantly higher isothermal compressibility in the region where $M<0$, near the base of the maxima ridge, compared to the compressibility values in the $M>0$ region.
-In the first case ($M<0$), the supercritical fluid exhibits gas-like behavior.
-In the second case ($M>0$), it behaves like a liquid.
-Far beyond the critical point ($\tau > 0.5$), there is no noticeable difference in $\kappa_T$ values around the effective chemical potential $M=0$.
-At $\tau = 0.2, P^*>2.30$ the compressibility maxima become almost undistinguished, they completely vanish at $\tau > 0.28$.
-\begin{figure}[h!] 
-	%	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
-	\centering \includegraphics[width=0.46\textwidth]{f3cc.pdf}
-	%\includegraphics[width=\column]{f2b.pdf}
-	%\includegraphics[width=\column]{f2c.pdf}
-	\vskip-3mm
-	\caption{The reduced thermal expansion $\alpha^*_P$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$.
-The plot focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\alpha^*_P$.
-	}
-	\label{fig3b}
-\end{figure}
-To identify the Widom line based on the locus of the isothermal compressibility maxima, we solve the equation
-\begin{equation}\label{eq:maximize}
-	\frac{\partial \kappa_T^*}{\partial M} = 0
-\end{equation}
-with respect to $M$.
-Substituting the solution of \eqref{eq:maximize} into the equation of state \eqref{eq:eosMT}, one locates the Widom line in the pressure-temperature plane (see Figure~\ref{fig4}, dashed line), expanding the phase diagram displayed in Figure~\ref{fig2} beyond the critical point.
-
-Similarly to the isothermal compressibility, similar arguments apply to the thermal expansion coefficient defined as
-\begin{equation}
-	\alpha_{P} = \frac{1}{V} \left(\frac{\partial V}{\partial T}\right)_{P,N}.
-\end{equation}
-Its reduced counterpart is calculated based on EoS~\eqref{eq:eosMT} by formula
-\begin{eqnarray}
-	\alpha^*_P & = & \frac{1}{T^*_c \rho^*}
-	\left[ 
-	-\left(\frac{\partial \rho^*}{\partial \tau}\right)_{\mu}
-	\right. \nonumber\\
-	&& + \left.\left(\frac{\partial \rho^*}{\partial \mu}\right)_{T}
-	\left(\frac{\partial P^*}{\partial \tau}\right)_{\mu}
-	\left(\frac{\partial P^*}{\partial \mu}\right)^{-1}_{T} 
-	\right].
-\end{eqnarray}
-Figure~\ref{fig3b} illustrates the dependence of $\alpha^*_{P}$ on $M$ at some given values of $\tau$, at $T>T_c$.
-The thermal expansion coefficient also develops a ridge of maxima in the supercritical phase, that almost vanishes at $\tau > 0.35, P^*>2.75$.
-To identify the Widom line based on the locus of the thermal expansion coefficient maxima, we solve the equation
-\begin{equation}\label{eq:maximize_alpha}
-	\frac{\partial \alpha_P^*}{\partial M} = 0
-\end{equation}
-with respect to $M$.
-Substituting the solution of \eqref{eq:maximize_alpha} into the equation of state \eqref{eq:eosMT}, one locates another Widom line in the pressure-temperature plane (see Figure~\ref{fig4}, dash-dotted line).
-This is the second option for the separation line between liquid-like and gas-like fluid.
-
-The third choice for the boundary between liquid-like and gas-like regions beyond CP is based on the condition $M=0$.
-Within the present approach, the condition $M=0$ holds along phase coexistence up to the critical point.
-Thus, the implementation of the condition $M=0$ above CP seems to be a natural continuation of the coexistence line into the supercritical region.
-On the other hand, taking into account the functional form of the grand partition function~\cite{KD16,KR09}, the condition of $M=0$ is equivalent to the absence of an external field in the Ising model.
-This analogy is interesting from the perspective of direct mapping between the Ising model and fluids~\cite{Kulinskii10jcp,BK11}.
-In Fig.~\ref{fig4} the separation line based on this condition $M=0$ is given by the dotted curve.
-In the critical region ($\tau<0.02$), beyond CP, the maxima locus of the response functions and the $M=0$ based line visually merge into a single line.
-
-\begin{figure}[h!]
-	\centering \includegraphics[width=0.5\textwidth]{f4b.pdf}
-	\vskip-3mm\caption{Complete pressure-temperature phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse}.
-Comparison of the Widom line (maxima locus of isothermal compressibility and thermal pressure) and the crossover between gas-like and liquid-like structures of the supercritical cell fluid based on combination of the equation of state~\eqref{eq:eosMT} and the condition $M = 0$.
-An inset in the right bottom corner focuses on a range of $\tau$ around its critical value $-0.01 < \tau < 0.05$.	}\label{fig4}
-\end{figure}
+\input{./paper_text.tex}
 
@@
 %\subsubsection{Порівняння обох ліній}
-
-\section{Conclusions}
-The supercritical region of the phase diagram is studied for the cell fluid model with the modified Morse potential.
-Three lines are found that can be considered as good candidates for a separation boundary between gas-like and liquid-like regions in the supercritical phase.
-The first line is the locus of the isothermal compressibility maxima.
-The second line is the locus of the thermal expansion coefficient maxima.
-The third line is the location of states where the condition $M=0$ holds true.
-This third option seems to be a natural extension of the coexistence line beyond the critical point, since the same condition is used both below and above $T_c$.
-A noticeable difference between two loci of thermodynamic response functions maxima and the line obtained from $M=0$ is that the latter one is extended over the phase diagram without any limits, while the first two have their termination points.
-Investigation of how physical quantities change their behavior when crossing these lines will be the subject of our future studies.
-
-\vskip3mm \textit{This work was supported by the National Research Foundation of Ukraine under the project No.~2023.03/0201.}
 

### 2024-10-30 paper_text.tex
add UJP styles

\section{Introduction}
A supercritical fluid is a state of matter that occurs when a substance is subjected to temperatures and pressures above its critical point (CP). Below CP the substance exhibits a liquid-gas coexistence. At CP, the distinction between the liquid and gas phases of a substance disappears, and beyond this point, the substance exists in a single continuous supercritical state. Despite the absence of first-order phase transitions above the critical point, fluids exhibit distinct liquid-like and gas-like behavior under supercritical conditions, and display a unique mix of properties. With densities comparable to liquids they possess low viscosity and high diffusivity like gases resulting in high solvent power. Moreover, the supercritical properties are instantly tunable by changing pressure and temperature. This is the reason why supercritical fluids have diverse applications across various industries \cite{KMLetal14}.

Although the traditional liquid-gas phase distinction does not apply in the supercritical region, scientists continue to search for crossover regions or definitive lines that distinguish liquid-like and gas-like behaviors in the phase diagram. Naturally, such boundaries originate from CP and may either vanish away from it or extend infinitely far. Among known approaches to this problem are the Widom line, which represents the locus of correlation length maxima~\cite{FW69}; the Frenkel line~\cite{BGMetal17}, focusing on the dynamics of molecular motion; and the percolation line~\cite{CKS01}, related to spatial structures and marking the threshold where clusters of fluid or density fluctuations form a continuous, system-spanning network.

In the present work, we focus on the Widom line \cite{XKBetal05,MS10,AKM17}, a concept based on thermodynamic response functions that represents a conditional boundary between liquid-like and gas-like regions in supercritical fluids. Thus, the Widom line \--- the locus of correlation length maxima emanating from CP~\cite{XKBetal05} \--- is treated as the continuation of the coexistence line above CP. Moreover, some thermodynamic response functions, such as isobaric heat capacity and isothermal compressibility, also exhibit maxima in the supercritical region. In the critical region beyond the critical point, the maxima lines of the correlation length and these response functions virtually merge into a single line~\cite{AKM17} and then diverge from each other as the temperature increases. Therefore, in the critical region, the Widom line is viewed as the locus of maxima of thermodynamic response functions~\cite{Proctor20}.

In our previous work~\cite{KD20}, the equation of state for a many-particle system interacting via a modified Morse potential was developed in the framework of a cell model. Recently~\cite{DKRP24,DKRP24arxiv}, the thermodynamic response functions, namely the isothermal compressibility, the thermal pressure coefficient, and the thermal expansion coefficient, have been calculated for the same model system. The isothermal compressibility and thermal expansion coefficient both develop a maximum in some region beyond critical point. On the one hand, such response function maxima, as mentioned above, are considered as good candidates for crossover lines between gas-like and liquid-like behaviors in the supercritical region. On the other hand, they are also considered as natural continuation of the coexistence line. Since our model allows for analytical expressions for many thermodynamic quantities~\cite{KD20,DKRP24}, it is straightforward to calculate and build such continuations. As a logical sequence of our previous works for the cell model with the modified Morse potential, in this paper we present results for three lines that can be cosidered as continuations of the coexistence line beyond the critical point in the pressure-temperature plane. The first one is the locus of the isothermal compressibility maxima. This line, as will be shown, terminates at some temperature above the critical one. The second line is the locus of the thermal expansion coefficient, which also terminates at some point. The third one is a line that is derived from the same condition that the first-order phase transition line below the critical point - the condition on the effective chemical potential $M=0$. This line can be extended into the supercritical region without any limits in temperature or pressure.

In Section~\ref{sec:eos}, we recall the form of the modified Morse potential, present the explicit expression for the equation of state, and illustrate pressure isotherms at temperatures below and above the critical one. In Section~\ref{sec:widom}, we develop two alternatives of the Widom line, as the maxima locus of the isothermal compressibility and as the locus of maxima of the thermal expansion coefficient. Then we introduce the supercritical crossover line obtained directly from the equation of state of the cell fluid model applying a condition on the effective chemical potential that is the same as is used below CP to obtain the coexistence line.

\section{\label{sec:eos} The equation of state}

In the present work, we utilize the results obtained in~\cite{KD20,DKRP24arxiv,DKRP24}. In~\cite{KD20}, the phase space of collective variables was used to calculate the grand partition function of a cell fluid model in the zero-mode approximation, leading to explicit forms of the equation of state (EoS) in both pressure-temperature-chemical potential and pressure-temperature-density terms.
These EoSs are applicable across a wide range of temperatures below and above the CP. Furthermore, in~\cite{DKRP24arxiv,DKRP24}, we have used these equations as the basis for calculating and graphically representing the thermodynamic response functions \--- specifically, the isothermal compressibility, thermal pressure coefficient, and thermal expansion \--- of the cell fluid model with the modified Morse potential in the supercritical region.

The modified Morse potential used to describe the interaction between particles is given by 
\begin{eqnarray}
	\label{def:mod_morse}
	U(r) & = & \varepsilon C_{H} \left[A {\rm e}^{-n_0(r-R_0)/\alpha} \right.
	\nonumber\\
	&& + \left. {\rm e}^{-\gamma(r - R_0)/\alpha} -2{\rm e}^{-(r-R_0)/\alpha}\right],
\end{eqnarray}
where $R_0$ is the coordinate of the potential minimum, $\alpha$ is an effective range of interaction, $\gamma$ and $n_0$ are parameters of the model. Other two constants $C_{H}$ and $A$ are expressed via $\gamma$ and $n_0$ as follows
\begin{equation}
	C_{H} = \frac{n_0}{n_0 + \gamma - 2}, \quad A = \frac{2 - \gamma}{n_0},
\end{equation}
where $\varepsilon$ is the depth of the potential well at $r=R_0$. Applying $\gamma=2$ reduces $U(r)$ to the ordinary Morse potential~\cite{Morse29}. For a more detailed discussion of such modified Morse potential, see Sections~1 and~2 in~\cite{KD20}, and Section~2 in~\cite{DKRP24arxiv}.

We use $\varepsilon$ and $v$ as natural units for energy and volume. Therefore, the following reduced quantities are defined: the reduced density $\rho^*=\rho v$, where $\rho = \langle N \rangle /V$; reduced temperature $T^* = k_{\rm B}T/\varepsilon$; reduced pressure $P^* = Pv/\varepsilon$; reduced isothermal compressibility $\kappa^*_{T} = \varepsilon \kappa_T/v$; and the reduced thermal expansion coefficient $\alpha^*_{P} = \varepsilon \alpha_{P} / k_{\rm B}$. In above expressions $V$ stands for volume, $T$ the temperature, $k_{\rm B}$ the Boltzmann constant, $P$ the pressure, $\kappa_T$ the isothermal compressibility, $\alpha_{P}$ the thermal expansion coefficient, $\langle N \rangle$ is the average number of particles, with averaging over the grand canonical distribution.

The equation of state obtained in~\cite{KD20} and rewritten in reduced quantities in~\cite{DKRP24,DKRP24arxiv} reads
\begin{eqnarray}\label{eq:eosMT}
	%\label{eq:eosPTM_reduced}
	P^* & = & (1 + \tau)T^*_c \bigg[ E_\mu(M, T) 
	\nonumber\\
	&& + M \bar \rho_0 + \frac{1}{2} d \bar \rho_0^2 - \frac{a_4}{24} \bar \rho_0^4
	\bigg].
\end{eqnarray}
Let us explicitly present the expressions entering the right-hand side of this equation.

First, the quantity $M$ depends linearly on the chemical potential
\begin{align}\label{chem_pot}
	&	M = \frac{\tilde\mu}{W(0)} + g_1 - \frac{g_3}{g_4} d - \frac{1}{6} \frac{g_3^3}{g_4^2}, \\
	&	\tilde\mu=\mu-\mu_0(1+\tau),
\end{align}
where $\mu_0$ is some positive constant, $\tau$ is the relative temperature $\tau = (T - T_c) / T_c$, $T_c$ is the critical temperature. We will call $M$ the effective chemical potential.

The quantity $W(0)$ is expressed via parameters of the potential~(\ref{def:mod_morse}) as follows
\begin{equation}
	W(0) = \Phi^{(r)}(0) \left[ B - 1 + \chi_0 + \tau (\chi_0 + A_\gamma) \right],
\end{equation}
where
\begin{align*} 
	& B = 2 \gamma^3 e^{(1-\gamma)R_0/\alpha},
	\nonumber \\
	& A_\gamma = A e^{(n_0-\gamma)R_0/\alpha} \left( \gamma / n_0\right)^3, 
\end{align*}
and $\Phi^{(r)}(0)$ is the Fourier transform of the repulsive part of the potential at $\abs{\vb k}=0$
\begin{equation*}
	\Phi^{(r)}(0) = \varepsilon C_H 8\pi {\rm e}^{\gamma R_0/\alpha} \left(\frac{\alpha}{\gamma R_0}\right)^3.
\end{equation*}

The parameter $\chi_0$ is used in~\cite{KD20} to single out a contribution in the Fourier transform of the potential that is treated as a reference system defined in the reciprocal space, and is selected as $\chi_0 = 0.07$~\cite{KD20}.

The coefficients $g_n$ are given by the formulas:
\begin{align}
	& g_0 = \ln T_0, \qquad g_1 = T_1/T_0, \qquad g_2 = T_2/T_0 - g_1^2,  
	\nonumber \\
	& g_3 = T_3/T_0 - g_1^3 - 3g_1 g_2, 
	\nonumber\\
	& g_4 = T_4/T_0 - g_1^4 - 6 g_1^2 g_2 - 4 g_1 g_3 - 3 g_2^2, 
\end{align}
where $T_n(p,\alpha^*)$ are the following special functions
\begin{equation}
	T_n(p,\alpha^*) = \sum_{m=0}^{\infty} \frac{(\alpha^*)^m}{m!} m^n {\rm e}^{-pm^2}.
\end{equation}
Here $\alpha^*=v e^{\beta_c\mu_0}$, where $\beta_c$ is the critical value of the inverse temperature $\beta = (k_{\rm B}T)^{-1}$, and the parameter $p$ has the form
\begin{equation}
	p = \frac{1}{2T^*_c} \frac{\Phi^{(r)}(0)}{\varepsilon} [\chi_0 + A_\gamma].
\end{equation} 
The critical temperature found in~\cite{KD20} is $T_c^* = 4.995.$

Since $p$ is independent of temperature, the coefficients $g_n$ are also independent of temperature. The numerical values for other coefficients used in this paper are the same as those in~\cite[eqs.~(5), (23), and~(24)]{KD20}:
\begin{eqnarray}
	\label{params}
	\chi_0 = 0.07, & \quad \gamma = 1.65, \nonumber\\
	n_0 = 1.521, & \quad R_0/\alpha = 2.9544, \nonumber\\
	\alpha^* = 5.0 & \quad p = 1.0.
\end{eqnarray}

The quantity $d$ entering equations~\eqref{eq:eosMT} and~\eqref{chem_pot} is a function of temperature
\begin{equation}
	\label{def:D0}
	d = g_2 - \frac{1}{2} \frac{g_3^2}{g_4} - \frac{1}{\beta W(0)}.
\end{equation}
The condition $d = 0$ defines the critical temperature~\cite{KD20}
\begin{equation}
	T^*_c = \left(g_2 - \frac{1}{2} \frac{g_3^2}{g_4} \right) (B - 1 + \chi_0) \frac{\Phi^{(r)}(0)}{\varepsilon}.
\end{equation}  
The function $E_\mu(M, T)$ from the equation \eqref{eq:eosMT} is given by
\begin{eqnarray}\label{eq:E_mu}
	E_\mu (M, T) & = & - \frac{\ln (2\pi \beta W(0))}{2 N_v}  
	\nonumber\\
	&& +  g_0 - \frac{\beta W(0)}{2} 
	\left(\frac{\tilde\mu}{W(0)} \right)^{2} 
	\nonumber\\
	&& - \frac{g_3}{g_4} {M} \! - \frac{g_3^2}{2 g_4^2}  d - \frac{1}{24} \frac{g_3^4}{g_4^3}. 
\end{eqnarray}
Here the quantity $N_v$ defines the number of cubic cells in volume $V$ for the initial model.
In the thermodynamic limit, $N_v \to \infty$, and thus, the first term can be neglected. The term $\tilde{\mu}/W(0)$ can be expressed in terms of $M$ using~\eqref{chem_pot}. The temperature and the inverse temperature can always be expressed in terms of the reduced temperature and a corresponding critical value:
\begin{equation*}
	T = T_c(1+\tau), \quad \beta = \beta_c (1 + \tau)^{-1}.
\end{equation*}

The quantity $\bar{\rho}_0$ is a solution to the following cubic equation
\begin{equation}\label{eq:ro_M}
	M + d \bar\rho_0 + \frac{g_4}{6} \bar\rho_0^3 = 0.
\end{equation}
For any $\tau > 0$, the latter equation has one real root
\begin{equation}\label{eq:ro_MT}
	\bar \rho_0 = \left(- \frac{3 M}{g_4} + \sqrt{Q_t}\right)^{1/3} - \left(  \frac{3 M}{g_4} + \sqrt{Q_t} \right)^{1/3},
\end{equation}
where
\begin{equation}
	Q_t = \left(  \frac{2d}{g_4}\right)^3 + \left( -\frac{3 M}{g_4}\right)^2, \qquad g_4<0.
\end{equation}
Thus, $\bar{\rho}_0$ is a function of the temperature and the chemical potential.

\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{f1.pdf} 
	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the effective chemical potential $M$. Black lines correspond to $T \geq T_c$. Grey lines correspond to $T < T_c$. 
	}\label{fig1}
\end{figure}
Figure~\ref{fig1} illustrates the relationship between the reduced pressure $P^*$ and the effective chemical potential $M$ for various relative temperature values $\tau$. Below the critical temperature ($\tau < 0$), each іsotherm contains a kink (three bottom lines in Figure~\ref{fig1}). This form indicates a first-order phase transition in the system. Within our approach~\cite{KD20} at $T<T_c$, negative values of the effective chemical potential $M$ imply the system is in a gaseous state. The condition $M=0$ is a point of phase coexistence. Positive values of the effective chemical potential correspond to the liquid state. In both cases, the pressure tends to increase with $M$, but the slope of the liquid isotherm is much steeper. At the critical temperature, the pressure is a smooth curve (black dotted line in Figure~\ref{fig1}) with an inflection point at $M=0$. Thus, at the critical point coordinates are defined by $M=0, \tau=0$, $P^*_c = 1.606$. Beyond the critical temperature, the pressure behaves as a monotonically increasing function, and no phase transitions are observed. Summing up, below the critical temperature, $M=0$ is the condition for phase coexistence. Substituting this condition into the EoS~\eqref{eq:eosMT}, we obtain a phase diagram in pressure-temperature plane (see Figure~\ref{fig2}). We will use the same condition to obtain a continuation of the coexistence line into the supercritical region. 
\begin{figure}[t!]
	\centering
	\includegraphics[width=0.45\textwidth]{f2.pdf} 
	\vskip-3mm\caption{Phase diagram in the $P^*-\tau$ plane for the cell fluid model with the interaction potential \eqref{def:mod_morse}. The result is based on the equation of state~\eqref{eq:eosMT}. 
	}\label{fig2}
\end{figure}

\section{\label{sec:widom} Crossover lines between gas-like and liquid-like behavior of the supercritical fluid.}

In the supercritical region, substances no longer behave distinctly as either liquids or gases but exist as a single, continuous phase without a sharp liquid-gas boundary. However, fluids can still exhibit liquid-like and gas-like behaviors beyond the critical temperature. A way to represent a conditional boundary between these regions in supercritical fluids is the Widom line \--- a continuation of the coexistence line based on the locus of maxima of certain thermodynamic response functions~\cite{XKBetal05,MS10,SBGetal10}. This occurs because near CP, the system experiences significant fluctuations in density over increasingly large spatial scales. Although, the equation of state~\eqref{eq:eosMT} is derived in the mean-field-like approximation, we established in~\cite{DKRP24arxiv,DKRP24} that the isothermal compressibility of the cell fluid model diverges at CP and consequently displays maxima in the supercritical region (see Figures~\ref{fig3}). This suggests the possibility of deriving the Widom line and expanding the phase diagram in Figure~\ref{fig2} beyond CP.
In what follows we consider this in more detail.

The isothermal compressibility $\kappa_T$ is defined by
\begin{equation}
	\label{def:isotherm_compres}
	\kappa_T = -\frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_{T, N}.
\end{equation}
In \cite{DKRP24arxiv,DKRP24} we derived expression for the reduced isothermal compressibility $\kappa^*_T \equiv \dfrac{\varepsilon \kappa_T}{v}$  in terms of thermodynamic derivatives suitable for the reduced form of the equation of state in pressure-temperture-chemical potential terms \eqref{eq:eosMT} 
It is expressed in terms of the reduced pressure $P^*$ and the reduced particle number density $\rho^*$ as follows
\begin{eqnarray}
	\label{eq:kappa_star_m1}
	\kappa^*_T & = & \frac{\varepsilon}{{\rho^*}^2} \left(\frac{\partial \rho^*}{\partial \mu}\right)_T
	\\
	\label{eq:kappa_star_m}
	& = & \frac{\varepsilon^2}{{\rho^*}^2} \left(\frac{\partial^2 P^*}{\partial \mu^2}\right)_T.
\end{eqnarray}
The reduced particle number density $\rho^*$ is found by 
\begin{equation}
	\rho^* \equiv \frac{\langle N \rangle}{V} v = \varepsilon\left(\frac{\partial P^*}{\partial \mu}\right)_{T,V}.
\end{equation}
Taking explicit derivatives, we arrive at
\begin{equation}\label{eq:density}
	\rho^* = \rho^*_c - M + \frac{ \bar \rho_0}{\beta W(0)}.
\end{equation}
The quantity $\rho^*_c$ in the equation \eqref{eq:density} is the critical density~\cite{KD20,DKRP24arxiv,DKRP24}
\begin{eqnarray}\label{eq:crit_dens}
	\rho^*_c & = &  g_1 - \frac{g_2 g_3}{g_4} + \frac{g_3^3}{3g_4^2}.
\end{eqnarray}
Its numerical value for parameters~\eqref{params} is $\rho^*_c = 0.978$.
Finally, the isothermal compressibility takes on the explicit form  
\begin{eqnarray}
	\kappa_T^* & = & \frac{\epsilon}{\rho^{*2} W(0)} \left\{   \frac{Q_t^{-1/2}}{\beta W(0) g_4} \left[ \bar \rho_0 - 2 \left( - \frac{3M}{g_4} + \sqrt{Q_t} \right)^{\frac{1}{3}} \right] -1 \right\}.
\end{eqnarray}
\begin{figure}[h!] 
	%	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
	\centering \includegraphics[width=0.46\textwidth]{f3bb.pdf}
	%\includegraphics[width=\column]{f2b.pdf}
	%\includegraphics[width=\column]{f2c.pdf}
	\vskip-3mm
	\caption{The reduced isothermal compressibility $\kappa^*_T$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$. The plot focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\kappa^*_T$.
	}
	\label{fig3}
\end{figure}
Behavior of the reduced isothermal compressibility versus effective chemical potential at various fixed values of the reduced temperature is represented graphically in Figure~\ref{fig3} (narrow range around $M=0$). Right at the critical point the compressibility diverges. As a result, there is still a "memory" of the critical point, where $\kappa_T$ reaches local maxima. However, once the system moves further into the supercritical region, these maxima become less pronounced, and the compressibility decreases. 
Below the critical temperature, negative values of $M$ correspond to gaseous densities, while positive values of $M$ represent liquid densities. Above the critical point, the fluid becomes continuous. However, Figure~\ref{fig3} clearly shows significantly higher isothermal compressibility in the region where $M<0$, near the base of the maxima ridge, compared to the compressibility values in the $M>0$ region. In the first case ($M<0$), the supercritical fluid exhibits gas-like behavior. In the second case ($M>0$), it behaves like a liquid. Far beyond the critical point ($\tau > 0.5$), there is no noticeable difference in $\kappa_T$ values around the effective chemical potential $M=0$. At $\tau = 0.2, P^*>2.30$ the compressibility maxima become almost undistinguished, they completely vanish at $\tau > 0.28$. 
\begin{figure}[h!] 
	%	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
	\centering \includegraphics[width=0.46\textwidth]{f3cc.pdf}
	%\includegraphics[width=\column]{f2b.pdf}
	%\includegraphics[width=\column]{f2c.pdf}
	\vskip-3mm
	\caption{The reduced thermal expansion $\alpha^*_P$ as a function of the effective chemical potential $M$ for different temperatures $\tau = (T - T_c)/T_c$ at $T > T_c$. The plot focuses on a range of $M$ around its critical value $0$, thin dashed black line displays maxima locus of $\alpha^*_P$.
	}
	\label{fig3b}
\end{figure}
To identify the Widom line based on the locus of the isothermal compressibility maxima, we solve the equation
\begin{equation}\label{eq:maximize}
	\frac{\partial \kappa_T^*}{\partial M} = 0
\end{equation}
with respect to $M$.
Substituting the solution of \eqref{eq:maximize} into the equation of state \eqref{eq:eosMT}, one locates the Widom line in the pressure-temperature plane (see Figure~\ref{fig4}, dashed line), expanding the phase diagram displayed in Figure~\ref{fig2} beyond the critical point.

Similarly to the isothermal compressibility, similar arguments apply to the thermal expansion coefficient defined as
\begin{equation}
	\alpha_{P} = \frac{1}{V} \left(\frac{\partial V}{\partial T}\right)_{P,N}.
\end{equation}
Its reduced counterpart is calculated based on EoS~\eqref{eq:eosMT} by formula
\begin{eqnarray}
	\alpha^*_P & = & \frac{1}{T^*_c \rho^*}
	\left[ 
	-\left(\frac{\partial \rho^*}{\partial \tau}\right)_{\mu}
	\right. \nonumber\\
	&& + \left.\left(\frac{\partial \rho^*}{\partial \mu}\right)_{T}
	\left(\frac{\partial P^*}{\partial \tau}\right)_{\mu}
	\left(\frac{\partial P^*}{\partial \mu}\right)^{-1}_{T} 
	\right].
\end{eqnarray}
Figure~\ref{fig3b} illustrates the dependence of $\alpha^*_{P}$ on $M$ at some given values of $\tau$, at $T>T_c$. The thermal expansion coefficient also develops a ridge of maxima in the supercritical phase, that almost vanishes at $\tau > 0.35, P^*>2.75$.
To identify the Widom line based on the locus of the thermal expansion coefficient maxima, we solve the equation
\begin{equation}\label{eq:maximize_alpha}
	\frac{\partial \alpha_P^*}{\partial M} = 0
\end{equation}
with respect to $M$.
Substituting the solution of \eqref{eq:maximize_alpha} into the equation of state \eqref{eq:eosMT}, one locates another Widom line in the pressure-temperature plane (see Figure~\ref{fig4}, dash-dotted line). This is the second option for the separation line between liquid-like and gas-like fluid.

The third choice for the boundary between liquid-like and gas-like regions beyond CP is based on the condition $M=0$. Within the present approach, the condition $M=0$ holds along phase coexistence up to the critical point. Thus, the implementation of the condition $M=0$ above CP seems to be a natural continuation of the coexistence line into the supercritical region. On the other hand, taking into account the functional form of the grand partition function~\cite{KD16,KR09}, the condition of $M=0$ is equivalent to the absence of an external field in the Ising model. This analogy is interesting from the perspective of direct mapping between the Ising model and fluids~\cite{Kulinskii10jcp,BK11}. In Fig.~\ref{fig4} the separation line based on this condition $M=0$ is given by the dotted curve. In the critical region ($\tau<0.02$), beyond CP, the maxima locus of the response functions and the $M=0$ based line visually merge into a single line.

\begin{figure}[h!]
	\centering \includegraphics[width=0.5\textwidth]{f4b.pdf}
	\vskip-3mm\caption{Complete pressure-temperature phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse}. Comparison of the Widom line (maxima locus of isothermal compressibility and thermal pressure) and the crossover between gas-like and liquid-like structures of the supercritical cell fluid based on combination of the equation of state~\eqref{eq:eosMT} and the condition $M = 0$. An inset in the right bottom corner focuses on a range of $\tau$ around its critical value $-0.01 < \tau < 0.05$.	}\label{fig4}
\end{figure}

\section{Conclusions}
The supercritical region of the phase diagram is studied for the cell fluid model with the modified Morse potential. Three lines are found that can be considered as good candidates for a separation boundary between gas-like and liquid-like regions in the supercritical phase. The first line is the locus of the isothermal compressibility maxima. The second line is the locus of the thermal expansion coefficient maxima. The third line is the location of states where the condition $M=0$ holds true. This third option seems to be a natural extension of the coexistence line beyond the critical point, since the same condition is used both below and above $T_c$. A noticeable difference between two loci of thermodynamic response functions maxima and the line obtained from $M=0$ is that the latter one is extended over the phase diagram without any limits, while the first two have their termination points. Investigation of how physical quantities change their behavior when crossing these lines will be the subject of our future studies.

\vskip3mm \textit{This work was supported by the National Research Foundation of Ukraine under the project No.~2023.03/0201.}

### 2024-10-30 kozlovskii_dobush_romanik_ujp_UTF8.tex
brushed for UJP, add ukrainian parts, add bibliography as bibitems, figures width=column

@@
 	{Supercritical crossover lines in a cell fluid model}%
+	\author{O.A.~Dobush}
 	\author{M.P.~Kozlovskii}%1
-	\author{O.A.~Dobush}
 	\author{R.V.~Romanik}
@@
 		A cell fluid model with a modified Morse potential is studied.
-The supercritical states are considered with respect to a possibility to build a separation boundary between liquid-like and gas-like bahaviors.
+The supercritical states are considered with respect to the possibility of building a separation boundary between liquid-like and gas-like behaviors.
 Three different lines are calculated that can be used for this purpose: the locus of the isothermal compressibility maxima, the locus of the thermal expansion coefficient maxima, and the line where the effective chemical potential is zero, $M=0$.
@@
 
-	\keywords{Cell model, Morse potential, Thermodynamic response functions}
+	\keywords{Cell fluid model, Morse potential, Supercritical region, Widom line}
 
@@
 	\input{./paper_text.tex}
-	\bibliographystyle{gost2008}
-	\bibliography{extracted}
+	%\bibliographystyle{gost2008}
+	%\bibliography{extracted}
 	%\bibliography{extracted}
@@
 	\vspace*{-5mm}\rezume{%Резюме укр. мовою
-		М.П.~Козловський, О.А.~Добуш, Р.В.~Романік}{Надкритичні лінії переходу в комірковій моделі плину} {В роботі проведено... .}{\textit{К\,л\,ю\,ч\,о\,в\,і\,
-			с\,л\,о\,в\,а:} коміркова модель, потенціал Морзе, термодинамічні функції відгуку.}
+		М.П.~Козловський, О.А.~Добуш, Р.В.~Романік}{Надкритичні лінії переходу в комірковій моделі плину}{Продовжується дослідження коміркової моделі плину з модифікованим потенціалом Морзе.
+Надкритичні стани розглядаються з огляду на можливість побудови межі розділення між рідиноподібними та газоподібними поведінками.
+Розраховано три різні лінії, які можна використовувати для цієї мети: лінія максимумів ізотермічної стисливості, лінія максимумів коефіцієнта теплового розширення та лінія, де ефективний хімічний потенціал дорівнює нулю, $M=0$.
+За симетрією функціоналів статистичних сум умова $M=0$ у рідинах аналогічна відсутності зовнішнього поля в моделі Ізінга.}{\textit{К\,л\,ю\,ч\,о\,в\,і\, с\,л\,о\,в\,а:} коміркова модель, потенціал Морзе, надкритична область, лінія Відома.}
 

### 2024-10-30 paper_text.tex
brushed for UJP, add ukrainian parts, add bibliography as bibitems, figures width=column

@@
 	\centering
-	\includegraphics[width=0.45\textwidth]{f1.pdf} 
+	\includegraphics[width=\column]{f1.pdf} 
 	\vskip-3mm\caption{Isotherms of the reduced pressure $P^*$ as a function of the effective chemical potential $M$. Black lines correspond to $T \geq T_c$. Grey lines correspond to $T < T_c$. 
@@
 	\centering
-	\includegraphics[width=0.45\textwidth]{f2.pdf} 
+	\includegraphics[width=\column]{f2.pdf} 
 	\vskip-3mm\caption{Phase diagram in the $P^*-\tau$ plane for the cell fluid model with the interaction potential \eqref{def:mod_morse}. The result is based on the equation of state~\eqref{eq:eosMT}. 
@@
 \begin{figure}[h!] 
-	%	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
-	\centering \includegraphics[width=0.46\textwidth]{f3bb.pdf}
-	%\includegraphics[width=\column]{f2b.pdf}
-	%\includegraphics[width=\column]{f2c.pdf}
+	\centering \includegraphics[width=\column]{f3bb.pdf}
 	\vskip-3mm
@@
 \begin{figure}[h!] 
-	%	\includegraphics[width=0.45\textwidth]{f3aa.pdf}
-	\centering \includegraphics[width=0.46\textwidth]{f3cc.pdf}
-	%\includegraphics[width=\column]{f2b.pdf}
-	%\includegraphics[width=\column]{f2c.pdf}
+	\centering \includegraphics[width=\column]{f3cc.pdf}
 	\vskip-3mm
@@
 \begin{figure}[h!]
-	\centering \includegraphics[width=0.5\textwidth]{f4b.pdf}
+	\centering \includegraphics[width=\column]{f4b.pdf}
 	\vskip-3mm\caption{Complete pressure-temperature phase diagram of the cell fluid model with the interaction potential \eqref{def:mod_morse}.
@@
 \vskip3mm \textit{This work was supported by the National Research Foundation of Ukraine under the project No.~2023.03/0201.}
+
+\begin{thebibliography}{99}
+
+	%1
+	\bibitem{KMLetal14} {\v{Z}}.~Knez, E.~Marko{\v{c}}i{\v{c}}, M.~Leitgeb, M.~Primo{\v{z}}i{\v{c}}, M.~Knez Hrn{\v{c}}i{\v{c}}, M.~{\v{S}}kerget.
+	\href{https://doi.org/10.1016/j.energy.2014.07.044}{Industrial applications of supercritical fluids: A review}. \textit{Energy} \textbf{77} 235 (2014) [doi: 10.1016/j.energy.2014.07.044]
+	%2
+	\bibitem{FW69} M.E.~Fisher, B.~Widom.
+	\href{https://doi.org/10.1063/1.1671624}{Decay of Correlations in Linear Systems}. \textit{J. Chem. Phys.} \textbf{50} 3756 (1969) [doi: 10.1063/1.1671624]
+	%3
+	\bibitem{BGMetal17} T.~Bryk, F.A.~Gorelli, I.~Mryglod, G.~Ruocco, M.~Santoro, T.~Scopigno.
+	\href{https://doi.org/10.1021/acs.jpclett.7b02176}{Behavior of Supercritical Fluids across the “Frenkel Line”}. \textit{J. Chem. Phys. Lett.} \textbf{8} 4995 (2017) [doi: 10.1021/acs.jpclett.7b02176]
+	%4
+	\bibitem{CKS01} X.~Campi, H.~Krivine, N.~Sator.
+	\href{https://doi.org/10.1016/S0378-4371(01)00158-3}{Percolation line of self-bound clusters in supercritical fluids}. \textit{Physica A} \textbf{296} 24 (2001) [doi: 10.1016/S0378-4371(01)00158-3] 
+	%5
+	\bibitem{XKBetal05} L.~Xu, P.~Kumar, S.V.~Buldyrev, S.-H.~Chen, P.H.~Poole, F.~Sciortino, H.E.~Stanley.
+	\href{https://doi.org/10.1073/pnas.0507870102}{Relation between the Widom line and the dynamic crossover in systems with a liquid-liquid phase transition}.
+\textit{PNAS} \textbf{102} 16558 (2005) [doi: 10.1073/pnas.0507870102]
+	%6
+	\bibitem{MS10} P.~McMillan, H.~Stanley.
+	\href{https://doi.org/10.1038/nphys1711}{Going supercritical}. \textit{Nature Phys.} \textbf{6} 479 (2010) [doi: 10.1038/nphys1711]
+	%7
+	\bibitem{AKM17} S.~Artemenko, P.~Krijgsman, V.~Mazur. 
+	\href{https://doi.org/10.1016/j.molliq.2017.03.107}{The Widom line for supercritical fluids}. \textit{J. Mol. Liq.} \textbf{238} 122 (2017) [doi: 10.1016/j.molliq.2017.03.107]
+	%8
+	\bibitem{Proctor20} J.E.~Proctor.
+	\href{https://doi.org/10.1201/9780429491443}{The liquid and supercritical fluid states of matter}. (CRC Press, 2020)  [ISBN: 9780429491443]
+	%9
+	\bibitem{KD20} M.~Kozlovskii, O.~Dobush.
+	\href{https://ujp.bitp.kiev.ua/index.php/ujp/article/view/2019631}{Phase behavior of a cell fluid model with modified Morse potential.} \textit{Ukr.
+J.
+Phys.} \textbf{65}, 428 (2020) [doi: 10.15407/ujpe65.5.428]
+	%10
+	\bibitem{DKRP24} O.A.~Dobush, M.P.~Kozlovskii, R.V.~Romanik, I.V.~Pylyuk.
+	\href{https://doi.org/10.48550/arXiv.2409.09786}{Thermodynamic response functions in a cell fluid model}. \textit{Ukr. J. Phys} - in print [doi: 10.48550/arXiv.2409.09786]
+	%11
+	\bibitem{DKRP24arxiv} O.A.~Dobush, M.P.~Kozlovskii, R.V.~Romanik, I.V.~Pylyuk.
+	\href{https://doi.org/10.48550/arXiv.2409.09786}{Thermodynamic response functions in a cell fluid model}. \textit{arXiv preprint} 2409.09786 (2024) [doi: 10.48550/arXiv.2409.09786]
+	%12
+	\bibitem{Morse29} P.~M.~Morse. 
+	\href{https://link.aps.org/doi/10.1103/PhysRev.34.57}{Diatomic molecules according to the wave mechanics.
+II.
+Vibrational levels.} \textit{Phys.
+Rev.} \textbf{34}, 57 (1929) [doi: 10.1103/PhysRev.34.57]
+	%13
+	\bibitem{SBGetal10} G.G.~Simeoni, T.~Bryk, F.A.~Gorelli, M.~Krisch, G.~Ruocco, M.~Santoro, T.~Scopigno.
+	\href{https://doi.org/10.1038/nphys1683}{ The Widom line as the crossover between liquid-like and gas-like behaviour in supercritical fluids}.
+\textit{Nature Phys.} \textbf{6} 503 (2010) [doi: 10.1038/nphys1683]
+	%14
+	\bibitem{KozlovskiiDobush2016} M.~Kozlovskii, O.~Dobush.
+	\href{https://doi.org/j.molliq.2015.12.018}{Representation of the grand partition function of the cell model: The state equation in the mean-field approximation.} \textit{J.
+Mol.
+Liq.} \textbf{215}, 58 (2016) [doi: j.molliq.2015.12.018]
+	%15
+	\bibitem{KR09} M.P~Kozlovskii, R.V.~Romanik.
+	\href{https://doi.org/10.30970/jps.13.4007}{The equation of state of a three-dimensional {Ising}-like system}. \textit{Journal of Physical Studies} \textbf{13} 4007 (2009) [doi: 10.30970/jps.13.4007]
+	%16
+	\bibitem{Kulinskii10jcp} V.L.~Kulinskii.
+	\href{https://doi.org/10.1063/1.3457943}{Global isomorphism between the {Lennard-Jones} fluids and the {Ising} model}. \textit{J. Chem. Phys.} \textbf{133} 034121 (2010) [doi: 10.1063/1.3457943]
+	%17
+	\bibitem{BK11} L.A.~Bulavin, V.L.~Kulinskii.
+	\href{https://doi.org/10.1021/jp201872f}{Unified Picture for the Classical Laws of {B}atschinski and the Rectilinear Diameter for Molecular Fluids}.
+\textit{J.
+Phys.
+Chem.
+B} \textbf{115} 6061 (2011) [doi: 10.1021/jp201872f]
+
+\end{thebibliography}

### 2024-10-30 paper_text.tex
changed author order

@@
 Moreover, the supercritical properties are instantly tunable by changing pressure and temperature.
-This is the reason why supercritical fluids have diverse applications across various industries \cite{KMLetal14}.
+This is the reason why supercritical fluids have diverse applications across various industries~\cite{KMLetal14}.
 
@@
 
-In the present work, we focus on the Widom line \cite{XKBetal05,MS10,AKM17}, a concept based on thermodynamic response functions that represents a conditional boundary between liquid-like and gas-like regions in supercritical fluids.
+In the present work, we focus on the Widom line~\cite{XKBetal05,MS10,AKM17}, a concept based on thermodynamic response functions that represents a conditional boundary between liquid-like and gas-like regions in supercritical fluids.
 Thus, the Widom line \--- the locus of correlation length maxima emanating from CP~\cite{XKBetal05} \--- is treated as the continuation of the coexistence line above CP.

### 2024-11-07 dobush_kozlovskii_romanik_ujp_UTF8.tex
last version

--- 2024-10-30 kozlovskii_dobush_romanik_ujp_UTF8.tex
@@
 	\author{R.V.~Romanik}
-	%\author{I.V.~Pylyuk}
+
 	\affiliation{Institute for Condensed Matter Physics,\\ Nat. Acad. of Sci. of Ukraine}%інститут 1
@@
 	\email{romanik@icmp.lviv.ua}
-	%\email{piv@icmp.lviv.ua}
 
@@
 	\vspace*{-5mm}\rezume{%Резюме укр. мовою
-		М.П.~Козловський, О.А.~Добуш, Р.В.~Романік}{Надкритичні лінії переходу в комірковій моделі плину}{Продовжується дослідження коміркової моделі плину з модифікованим потенціалом Морзе.
+		О.А.~Добуш, М.П.~Козловський, Р.В.~Романік}{Надкритичні лінії переходу в комірковій моделі плину}{Продовжується дослідження коміркової моделі плину з модифікованим потенціалом Морзе.
 Надкритичні стани розглядаються з огляду на можливість побудови межі розділення між рідиноподібними та газоподібними поведінками.

### 2024-11-10 paper_text.tex
fix long equation
fix ref to KD16
add two refs for mixtures

@@
 \begin{eqnarray}
-	\kappa_T^* & = & \frac{\epsilon}{\rho^{*2} W(0)} \left\{   \frac{Q_t^{-1/2}}{\beta W(0) g_4} \left[ \bar \rho_0 - 2 \left( - \frac{3M}{g_4} + \sqrt{Q_t} \right)^{\frac{1}{3}} \right] -1 \right\}.
+	\kappa_T^* & = & \frac{\epsilon}{\rho^{*2} W(0)}
+	\left\{   \frac{Q_t^{-1/2}}{\beta W(0) g_4} \right. \times
+	\\ 
+	&&\left. \times \left[ \bar \rho_0 - 2 \left( - \frac{3M}{g_4} + \sqrt{Q_t} \right)^{\frac{1}{3}} \right] -1 \right\}.
+	\nonumber
 \end{eqnarray}
@@
 On the other hand, taking into account the functional form of the grand partition function~\cite{KD16,KR09}, the condition of $M=0$ is equivalent to the absence of an external field in the Ising model.
-This analogy is interesting from the perspective of direct mapping between the Ising model and fluids~\cite{Kulinskii10jcp,BK11}.
+This analogy is interesting from the perspective of direct mapping between the Ising model and fluids~\cite{Kulinskii10jcp,BK11}, as well as from the consideration of other systems belonging to the Ising model class of universality, for example mixtures in criticality~\cite{OBP97,KD20b}.
 In Fig.~\ref{fig4} the separation line based on this condition $M=0$ is given by the dotted curve.
@@
 	%14
-	\bibitem{KozlovskiiDobush2016} M.~Kozlovskii, O.~Dobush.
+	\bibitem{KD16} M.~Kozlovskii, O.~Dobush.
 	\href{https://doi.org/j.molliq.2015.12.018}{Representation of the grand partition function of the cell model: The state equation in the mean-field approximation.} \textit{J.
@@
 	\bibitem{BK11} L.A.~Bulavin, V.L.~Kulinskii.
-	\href{https://doi.org/10.1021/jp201872f}{Unified Picture for the Classical Laws of {B}atschinski and the Rectilinear Diameter for Molecular Fluids}.
+	\href{https://doi.org/10.1021/jp201872f}{Unified picture for the classical laws of {B}atschinski and the rectilinear diameter for molecular fluids}.
 \textit{J.
@@
 B} \textbf{115} 6061 (2011) [doi: 10.1021/jp201872f]
+	\bibitem{OBP97} A.~Oleinikova, L.~Bulavin, V.~Pipich.
+	\href{https://doi.org/10.1016/S0009-2614(97)00945-7}{Critical anomaly of shear viscosity in a mixture with an ionic impurity}.
+\textit{Chem.
+Phys.
+Lett.} \textbf{278} 121 (1997) [doi: 10.1016/S0009-2614(97)00945-7]
+	\bibitem{KD20b} M.P.~Kozlovskii, O.A.~Dobush.
+	\href{https://doi.org/10.15407/ujpe65.9.768}{Behavior of a binary asymmetric mixture of interacting particles in the supercritical region}.
+\textit{Ukr.
+J.
+Phys.} \textbf{65} 768 (2020) [doi: 10.15407/ujpe65.9.768]
 
